---
title: Data Storytelling Team Project - Baseball
author: Byeong-Hak Choe
institute: SUNY Geneseo
date: 2024-11-04
# toc: true
# toc-depth: 2
# toc-expand: true
# toc-title: Contents
# number-sections: false
# number-depth: 3

code-fold: false

execute:
  echo: true
  eval: true
  message: false
  warning: false
  fig-width: 9
  fig-height: 6
---
```{r}
#| include: false
library(tidyverse)
library(skimr)
library(broom)
library(hrbrthemes)

theme_set(theme_ipsum() +
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = element_text(size = rel(1.5) ),
                axis.title.y = element_text(size = rel(1.5) ))
          )

# Set global options for color-blind-friendly scales
# scale_colour_discrete <- function(...) scale_colour_viridis_d(...)
scale_colour_discrete <- function(...) scale_color_colorblind(...)
scale_fill_discrete <- function(...) scale_fill_colorblind(...)
```

# Data

The following lists data frames about MLB since 1985:

- `df_teams`: Yearly statistics and standings for MLB teams
- `df_battings`: Batting statistics
- `df_pitchings`: Pitching statistics
- `df_salaries`: Player salary data
- `df_salariesTeam`: Team salary data
- `df_postseasons`: Post season series information
- `df_players`: People table - Player names, DOB, and biographical info. 
  - This data.frame is to be used to get details about players listed in the `df_battings`, `df_pitchings`, and `df_salaries` where players are identified only by variable `playerID`.


## MLB Teams (`df_teams`)
- `df_teams`: Yearly statistics and standings for MLB teams
  - A data frame with 1128 observations on the 52 variables.

```{r}
df_teams <- read_csv("http://bcdanl.github.io/data/mlb_teams.csv")
```

```{r}
#| echo: false
rmarkdown::paged_table(df_teams)
```

<br>

`yearID`
Year

`lgID`
League; a factor with levels AA AL FL NL PL UA

`teamID`
Team; a factor

`franchID`
Franchise (links to TeamsFranchises table)

`divID`
Team's division; a factor with levels C E W

`Rank`
Position in final standings

`G`
Games played

`Ghome`
Games played at home

`W`
Wins

`L`
Losses

`DivWin`
Division Winner (Y or N)

`WCWin`
Wild Card Winner (Y or N)

`LgWin`
League Champion(Y or N)

`WSWin`
World Series Winner (Y or N)

`R`
Runs scored

`AB`
At bats

`H`
Hits by batters

`X1B`
Singles

`X2B`
Doubles

`X3B`
Triples

`HR`
Homeruns by batters

`TB` 
Total bases
`TB = X1B + 2*X2B + 3*X3B + 4*HR`

`BB`
Walks by batters

`SO`
Strikeouts by batters

`SB`
Stolen bases

`CS`
Caught stealing

`HBP`
Batters hit by pitch

`BB_HBP`
`BB_HBP = BB + HBP`

`RC`
Runs created `RC = (H + BB + HBP)*TB/(AB + BB + HBP)`

`SF`
Sacrifice flies

`RA`
Opponents runs scored

`ER`
Earned runs allowed

`ERA`
Earned run average

`CG`
Complete games

`SHO`
Shutouts

`SV`
Saves

`IPouts`
Outs Pitched (innings pitched x 3)

`HA`
Hits allowed

`HRA`
Homeruns allowed

`BBA`
Walks allowed

`SOA`
Strikeouts by pitchers

`E`
Errors

`DP`
Double Plays

`FP`
Fielding percentage

`name`
Team's full name

`park`
Name of team's home ballpark

`attendance`
Home attendance total

`BPF`
Three-year park factor for batters

`PPF`
Three-year park factor for pitchers

`teamIDBR`
Team ID used by Baseball Reference website

`teamIDlahman45`
Team ID used in Lahman database version 4.5

`teamIDretro`
Team ID used by Retrosheet

## MLB Batting (`df_battings`)
- `df_battings`: Batting statistics
  - A data frame with 113799 observations on the 28 variables.


```{r}
df_battings <- read_csv("http://bcdanl.github.io/data/mlb_battings.csv")
```

```{r}
#| echo: false
rmarkdown::paged_table(df_battings)
```

<br>

`playerID`
Player ID code

`yearID`
Year

`stint`
player's stint (order of appearances within a season)

`teamID`
Team; a factor

`lgID`
League; a factor with levels AA AL FL NL PL UA

`G`
Games: number of games in which a player played

`AB`
At Bats

`R`
Runs

`H`
Hits: times reached base because of a batted, fair ball without error by the defense

`X2B`
Singles

`X2B`
Doubles: hits on which the batter reached second base safely

`X3B`
Triples: hits on which the batter reached third base safely

`HR`
Homeruns

`TB` 
Total bases
`TB = X1B + 2*X2B + 3*X3B + 4*HR`

`RBI`
Runs Batted In

`SB`
Stolen Bases

`CS`
Caught Stealing

`BB`
Base on Balls

`SO`
Strikeouts

`IBB`
Intentional walks

`HBP`
Hit by pitch

`BB_HBP`
Representing all the times a player reaches base via walks (including intentional walks) and hit by pitches.
`BB_HBP = BB + HBP`

`RC`
Runs created `RC = (H + BB + HBP)*TB/(AB + BB + HBP)`

`SH`
Sacrifice hits

`SF`
Sacrifice flies

`GIDP`
Grounded into double plays

`Outs`
The total number of outs a player is responsible for during a season
`Outs = 0.982 * AB - H + GIDP + SF + SH + CS`
Multiply AB by 0.982 due to the fact that approximately 0.8% of all at bats result in an error

`OBP`
On-Base Percentage.
OBP measures how frequently a batter reaches base per plate appearance
`OBP = (H + BB + HBP) / (AB + BB + HBP + SF)`


## MLB Pitching (`df_pitchings`)

- `df_pitchings`: Pitching statistics
  - A data frame with 26384 observations on the 33 variables.

```{r}
df_pitchings <- read_csv("http://bcdanl.github.io/data/mlb_pitchings.csv")
```

```{r}
#| echo: false
rmarkdown::paged_table(df_pitchings)
```

<br>

`playerID`
Player ID code

`yearID`
Year

`stint`
player's stint (order of appearances within a season)

`teamID`
Team; a factor

`lgID`
League; a factor with levels AA AL FL NL PL UA

`W`
Wins

`L`
Losses

`G`
Games

`GS`
Games Started

`CG`
Complete Games

`SHO`
Shutouts

`SV`
Saves

`IPouts`
Outs Pitched (innings pitched x 3)

`H`
Hits

`BAOpp`
Batting average against (a measure of how effectively a pitcher prevents hitters from getting hits)
`BAOpp = H / (H + IPouts)`

 
`ER`
Earned Runs

`HR`
Homeruns

`BB`
Walks

`WHIP`
Walks plus Hits per Inning Pitched. 
The number of base runners a pitcher has allowed per inning pitched. 
It is a common indicator of a pitcher’s effectiveness.
`WHIP = (H + BB) * 3 / IPouts`

`SO`
Strikeouts

`BAOpp`
Opponent's Batting Average

`ERA`
Earned Run Average

`IBB`
Intentional Walks

`KperBB`
Strikeouts per Walk.
KperBB represents the ratio of strikeouts to walks, indicating a pitcher’s control and ability to dominate hitters.
`KperBB = SO/(BB - IBB)`

`WP`
Wild Pitches

`HBP`
Batters Hit By Pitch

`BK`
Balks

`BFP`
Batters faced by Pitcher

`GF`
Games Finished

`R`
Runs Allowed

`SH`
Sacrifices by opposing batters

`SF`
Sacrifice flies by opposing batters

`GIDP`
Grounded into double plays by opposing batter


## MLB Salaries (`df_salaries`)

- `df_salaries`: Player salary data from 1985 to 2016
  - A data frame with 26428 observations on the 5 variables.

```{r}
df_salaries <- read_csv("http://bcdanl.github.io/data/mlb_salaries.csv")
```

```{r}
#| echo: false
rmarkdown::paged_table(df_salaries)
```

<br>

`yearID`
Year

`teamID`
Team; a factor

`lgID`
League; a factor

`playerID`
Player ID code

`salary`
Salary

### Team Salaries (`df_salariesTeam`)
- `df_salariesTeam`: Team salary data from 1985 to 2016

```{r}
df_salariesTeam <- read_csv("http://bcdanl.github.io/data/mlb_payrolls.csv")
```

```{r}
#| echo: false
rmarkdown::paged_table(df_salariesTeam)
```

`payroll` 
Team payroll in million dollars

`WSWin`
Whether or not winning the World Series


<br>



## MLB Post Season (`df_postseasons`)
- `df_postseasons`: Post season series information
  - A data frame with 389 observations on the 9 variables.

```{r}
df_postseasons <- read_csv("http://bcdanl.github.io/data/mlb_postseasons.csv")
```

```{r}
#| echo: false
rmarkdown::paged_table(df_postseasons)
```

<br>

`yearID`
Year

`round`
Level of playoffs

`teamIDwinner`
Team ID of the team that won the series; a factor

`lgIDwinner`
League ID of the team that won the series; a factor with levels AL NL

`teamIDloser`
Team ID of the team that lost the series; a factor

`lgIDloser`
League ID of the team that lost the series; a factor with levels AL NL

`wins`
Wins by team that won the series

`losses`
Losses by team that won the series

`ties`
Tie games


## MLB Players (`df_players`)
- `df_players`: People table - Player names, DOB, and biographical info. 
  - A data frame with 21010 observations on the 26 variables.
  - This data.frame is to be used to get details about players listed in the `df_battings`, `df_pitchings`, and `df_salaries` where players are identified only by variable `playerID`.

```{r}
df_players <- read_csv("http://bcdanl.github.io/data/mlb_players.csv")
```

```{r}
#| echo: false
rmarkdown::paged_table(df_players)
```

<br>

`playerID`
A unique code assigned to each player. 
The `playerID` links the data in this data.frame with records on players in the other data.frames.

`birthYear`
Year player was born

`birthMonth`
Month player was born

`birthDay`
Day player was born

`birthCountry`
Country where player was born

`birthState`
State where player was born

`birthCity`
City where player was born

`deathYear`
Year player died

`deathMonth`
Month player died

`deathDay`
Day player died

`deathCountry`
Country where player died

`deathState`
State where player died

`deathCity`
City where player died

`nameFirst`
Player's first name

`nameLast`
Player's last name

`nameGiven`
Player's given name (typically first and middle)

`weight`
Player's weight in pounds

`height`
Player's height in inches

`bats`
a factor: Player's batting hand (left (L), right (R), or both (B))

`throws`
a factor: Player's throwing hand (left(L) or right(R))

`debut`
Date that player made first major league appearance

`finalGame`
Date that player made first major league appearance (blank if still active)

`retroID`
ID used by retrosheet, https://www.retrosheet.org/

`bbrefID`
ID used by Baseball Reference website, https://www.baseball-reference.com/

`birthDate`
Player's birthdate, in as.Date format

`deathDate`
Player's deathdate, in as.Date format



# Sabermetrics

Baseball analysis underwent a significant transformation with Bill James’s pioneering work in the late 1970s. Through his publication, the Bill James Baseball Abstract, James challenged the reliance on traditional metrics like batting average, runs, and RBIs for hitters, and wins, ERA, and strikeouts for pitchers. Instead, he introduced innovative metrics such as runs created, Pythagorean win percentage, and game score. He called this analytical framework sabermetrics, named after the Society for American Baseball Research (SABR). While initially dismissed by traditionalists as secondary to the expertise of seasoned “baseball people,” James’s methods have become foundational in modern baseball. Today, every Major League Baseball team employs analytics experts who utilize advanced mathematics, predictive modeling, and optimization techniques to assemble competitive teams and make in-game decisions aimed at maximizing success.

For many, the 2011 movie Moneyball provides the most accessible introduction to sports analytics. The film recounts how the Oakland A’s used data-driven insights during their 2002 season, despite skepticism from traditional scouts who mocked Bill James’s ideas. Since that era, analytics have not replaced scouts but have become an essential complement, with every team now leveraging sophisticated data analyses alongside traditional evaluations.

# Baseball Metrics
## Historical Metrics
For over a century, baseball players were evaluated using a limited set of straightforward offensive and defensive statistics. However, advancements in computing and the introduction of sabermetrics revolutionized how player performance is assessed, giving rise to more nuanced and meaningful metrics.

### Offensive Statistics

| Statistic                   | Definition                                                                          |
|-----------------------------|-------------------------------------------------------------------------------------|
| **Batting Average**         | Hits divided by at-bats (Hits ÷ At-Bats)                                            |
| **Home Runs**               | Total number of home runs hit by a player                                           |
| **Runs Batted In (RBI)**    | Number of runs a player has batted in                                               |
| **Walks**                   | Number of times a batter reaches first base via four balls (also known as Base on Balls) |
| **On Base Percentage (OBP)**| Times on base (hits + walks + hit by pitch) divided by plate appearances            |
| **Runs Scored**             | Number of times a player crosses home plate to score a run                          |
| **Slugging Percentage (SLG)**| Total bases divided by at-bats (Total Bases ÷ At-Bats)                             |

### Defensive Statistics

| Statistic                    | Definition                                                                          |
|------------------------------|-------------------------------------------------------------------------------------|
| **Wins**                     | Credited to the pitcher who was pitching when his team took the lead for good       |
| **Innings Pitched**          | Total number of innings a pitcher has pitched                                       |
| **Earned Run Average (ERA)** | Average number of earned runs a pitcher allows per nine innings pitched             |
| **Walks**                    | Number of batters a pitcher has allowed to reach first base via four balls          |
| **Strikeouts**               | Number of batters a pitcher has retired by striking them out                        |
| **WHIP**                     | Walks plus hits per inning pitched ((Walks + Hits) ÷ Innings Pitched)               |
| **Saves**                    | Credited to a pitcher who finishes a game for the winning team under certain conditions |
| **Fielding Percentage**      | Percentage of times a defensive player handles the ball without making an error     |

## Runs Created
One of Bill James’s original metrics, Runs Created ($RC$), estimates the number of runs a player or team contributes to scoring. The formula is:

$$
RC = \frac{(H + BB + HBP) \times TB}{AB + BB + HBP}
$$

Where:

-	$H$ = Hits
-	$BB$ = Walks (Base on Balls)
-	$HBP$ = Hit by Pitch
-	$TB$ = Total Bases
-	$AB$ = At-Bats

This metric helps evaluate a player or team’s offensive value. For example, if considering adding a player to a team, comparing their runs created to other candidates provides a data-driven basis for decision-making.