---
title: Data Storytelling Team Project - Basketball
author: Byeong-Hak Choe
institute: SUNY Geneseo
date: last-modified

toc: true
toc-depth: 2
toc-expand: true
toc-title: Contents
number-sections: false
number-depth: 3

code-fold: false

execute:
  echo: true
  eval: true
  message: false
  warning: false
  fig-width: 9
  fig-height: 6
---
```{r}
#| include: false
library(tidyverse)
library(skimr)
library(broom)
library(hrbrthemes)

theme_set(theme_ipsum() +
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = element_text(size = rel(1.5) ),
                axis.title.y = element_text(size = rel(1.5) ))
          )

# Set global options for color-blind-friendly scales
# scale_colour_discrete <- function(...) scale_colour_viridis_d(...)
scale_colour_discrete <- function(...) scale_color_colorblind(...)
scale_fill_discrete <- function(...) scale_fill_colorblind(...)
```
 
# Data
The following lists data frames about NBA for the 2022-23 season:

- `nba_games`: NBA Game Logs
- `nba_adv`: NBA Advanced Statistics
- `nba_ff`: NBA Four Factors---(1) Effective field goal percentage, (2) Turnovers committed per possession, (3) Offensive rebounding percentage, and (4) Free throw rate
- `nba_games_adv_ff`: Joined data.frame (Game Logs + Advanced Statistics + Four Factors)
  - `nba_games_adv_ff <- nba_games |> left_join(nba_adv) |> left_join(nba_ff)` by `c("team_id", "team_abbreviation", "game_id", "min")`
  - Turnovers committed per possession is included here.
- `nba_players`: NBA Player Statistics
  - While the above data.frames are team-level data, this `nba_players` data.frame is player-level data.




## NBA Game Logs (`nba_games`)

- `nba_games`: NBA Game Logs
  - A data frame with 2460 observations and 29 variables:

```{r}
nba_games <- read_csv("http://bcdanl.github.io/data/nba_games.csv")
```

```{r}
#| echo: false
rmarkdown::paged_table(nba_games)
```

<br>

`season_id`
Season in which the game was played (YYYY)

`team_id`
Team ID on nba.com

`team_abbreviation`
Team abbreviation on nba.com

`team_name`
Team name on nba.com

`game_id`
Game ID on nba.com

`game_date`
Date in YYYY-MM-DD

`matchup`
Team One vs. Team Two

`wl`
Win or Loss

`min`
Minutes in the game

`fgm`
Team Field goals made

`fga`
Team Field goal attempts


`fg_pct`
Team field goal percentage

`fg3m`
Team three point field goals made

`fg3a`
Team three point field goal attempts

`fg3_pct`
Team three point field goal percentage

`efg`
Effective field goal percentage
`efg = (fg2m + 1.5*fg3m)/(fga)`
It adjusts the traditional field goal percentage to account for the fact that three-point field goals are worth more than two-point field goals.


`ftm`
Team free throws made

`fta`
Team free throw attempts

`ftr`
Free throw rate (FTR). The ratio of free throw attempts (`fta`) to field goal attempts (`fga`) `ftr = fta / fga`

`ft_pct`
Team free throw percentage

`oreb`
Team offensive rebounds

`dreb`
Team defensive rebounds

`reb`
Team total rebounds

`ast`
Team total assists

`stl`
Team total steals

`blk`
Team total blocks

`tov`
Team total turnovers

`pf`
Team total personal fouls

`pts`
Team total points scored

`plus_minus`
Margin of game as Team score minus Opponent score

`video_available`
Logical if game video exists


<br>

## NBA Advanced Statistics (`nba_adv`)
- `nba_adv`: NBA Advanced Statistics
  - A data frame with 2460 observations and 29 variables

```{r}
nba_adv <- read_csv("http://bcdanl.github.io/data/nba_adv.csv")
```

```{r}
#| echo: false
rmarkdown::paged_table(nba_adv)
```

<br>

`game_id`
Game ID on nba.com

`team_id`
Team ID on nba.com

`team_name`
Team name

`team_abbreviation`
Team abbreviation

`team_city`
Team city

`min`
Total minutes in the game

`e_off_rating`
Offensive rating (effective ?)

`off_rating`
Offensive rating

`e_def_rating`
Defensive rating (effective ?)

`def_rating`
Defensive rating

`e_net_rating`
Net rating (effective ?)

`net_rating`
Net rating

`ast_pct`
Assist percentage

`ast_tov`
Assists to turnover ratio

`ast_ratio`
Assist ratio

`oreb_pct`
Offensive rebound percentage

`dreb_pct`
Defensive rebound percentage

`reb_pct`
Total rebound percentage

`e_tm_tov_pct`
Turnover percentage (effective?)

`tm_tov_pct`
Turnover percentage

`efg_pct`
Effective field goal percentage

`ts_pct`
True Shooting percentage

`usg_pct`
Usage percentage

`e_usg_pct`
Usage percentage (effective?)

`e_pace`
Pace (effective ?)

`pace`
Pace

`pace_per40`
Pace per forty minutes

`poss`
Team possessions in game

`pie`
Player impact estimate



<br>


## NBA Four Factors (`nba_ff`)

- `nba_ff`: NBA Team-level Four Factors---(1) Effective field goal percentage, (2) Turnovers committed per possession, (3) Offensive rebounding percentage, and (4) Free throw rate
  - A data frame with 2460 observations and 14 variables

```{r}
nba_ff <- read_csv("http://bcdanl.github.io/data/nba_ff.csv")
```

```{r}
#| echo: false
rmarkdown::paged_table(nba_ff)
```

<br>

`game_id`
Game ID on nba.com

`team_id`
Team ID on nba.com

`team_name`
Team name on nba.com

`team_abbreviation`
Team abbreviation on nba.com

`team_city`
Team city

`min`
Minutes in the game

`efg_pct`
Effective field goal percentage

`fta_rate`
Free throw rate of team

`tm_tov_pct`
Turnover percentage of opponent

`oreb_pct`
Offensive rebound percentage of opponent

`opp_efg_pct`
Effective field goal percentage of opponent

`opp_fta_rate`
Free throw rate of opponent

`opp_tov_pct`
Turnover percentage of opponent

`opp_oreb_pct`
Offensive rebound percentage of opponent


<br>

## NBA Game Logs + Advanced Statistics + Four Factors (`nba_games_adv_ff`)

- `nba_games_adv_ff` is the data.frame that contains all the information in the `nba_games`, `nba_adv`, and `nba_ff` data.frames.
  - Turnovers committed per possession is included here.

```{r}
nba_games_adv_ff <- read_csv("http://bcdanl.github.io/data/nba_games_adv_ff.csv")
```

```{r}
#| echo: false
rmarkdown::paged_table(nba_games_adv_ff)
```

`tpp` 
Turnovers committed per possession (TPP). The number of turnovers committed relative to the number of possessions the team had during the game.
`tpp = tov / poss`. 

- `tov` comes from game log data.frame 
- `poss` comes from advanced statistics data.frame


## NBA Player Statistics (`nba_players`)

- `nba_players`: NBA Player Statistics
  - A data frame with 467 observations and 51 variables


```{r}
nba_players <- read_csv("http://bcdanl.github.io/data/nba_players.csv")
```

```{r}
#| echo: false
rmarkdown::paged_table(nba_players)
```

<br>

`PlayerName`
Player Name

`Salary`
Salary in US Dollars

`Position`
Position of Player

`Age`
Age

`Team`
Team(s) Played for in 2022-23 Season

`GP`
Games Played

`GS`
Games Started

`MP`
Minutes Per Game

`FG`
Field Goals Made Per Game

`FGA`
Field Goal Attempts Per Game

`FG_pct` 
Field Goal Percentage

`3P`
Three Point Field Goals Made Per Game

`3PA`
Three Point Field Goal Attempts Per Game

`3P_pct` 
Three Point Percentage

`2P`
Two Point Field Goals Made Per Game

`2PA`
Two Point Field Goal Attempts Per Game

`eFG_pct` 
Effective Field Goal Percentage

`FT`
Free Throws Made Per Game

`FTA`
Free Throw Attempts Per Game

`FT_pct` 
Free Throw Percentage

`ORB`
Offensive Rebounds Per Game

`DRB`
Defensive Rebounds Per Game

`TRB`
Total Rebounds Per Game

`AST`
Assists Per Game

`STL`
Steals Per Game

`BLK`
Blocks Per Game

`TOV`
Turnovers Per Game

`PF`
Personal Fouls Per Game

``PTS``
Points Per Game

`TotalMinutes`
Total Minutes Played in the Season

``PER``
Player Efficiency Rating

`TS_pct` 
True Shooting Percentage

`3PAr`
Three Point Attempt Rate

`FTr`
Free Throw Rate

`ORB_pct` 
Offensive Rebound Percentage

`DRB_pct` 
Defensive Rebound Percentage

`TRB_pct` 
Total Rebound Percentage

`AST_pct` 
Assist Percentage

`STL_pct` 
Steal Percentage

`BLK_pct` 
Block Percentage

`TOV_pct` 
Turnover Percentage

`USG_pct` 
Usage Rate

`OWS`
Offensive Win Shares

`DWS`
Defensive Win Shares

`WS`
Win Shares

`WS_per_48`
Win Shares Per 48 Minutes

`OBPM`
Offensive Box Plus/Minus

`DBPM`
Defensive Box Plus/Minus

`BPM`
Box Plus/Minus
 
`VORP`
Value Over Replacement Player

<br><br>

# Basketball Metrics

## Offensive Statistics

| Statistic                  | Definition                                                                 |
|----------------------------|---------------------------------------------------------------------------|
| **Points Scored**          | Total points scored by a player                                           |
| **Minutes Played**         | Amount of game time a player is one of the five active players on the court|
| **Field Goal Percentage**  | Number of shots made divided by the number of shots attempted (FG% = Made ÷ Attempted) |
| **Free Throw Percentage**  | Number of free throws made divided by the number of free throws attempted (FT% = Made ÷ Attempted) |
| **Assists**                | Passes that directly result in a teammate scoring a field goal            |
| **Three-Point Shots**      | Shots made from behind the three-point line                               |

## Defensive Statistics

| Statistic                  | Definition                                                                 |
|----------------------------|---------------------------------------------------------------------------|
| **Defensive Rebounds**     | Catching the ball after an opponent's missed shot                         |
| **Blocks**                 | Swatting the ball away as it leaves the opponent's hand                   |
| **Steals**                 | Taking the ball away from an opponent without committing a foul           |
| **Fouls**                  | Breaking the rules of permissible conduct, usually involving physical contact during a shot |
| **Turnovers**              | Actions that cause the other team to gain possession, excluding fouls or missed shots (e.g., bad passes out of bounds) |
| **Plus/Minus**             | The net number of points your team scored versus allowed while you were in the game |


## Four Factors
The “four factors” concept, developed by Dean Oliver and detailed in his book Basketball on Paper (available at basketballonpaper.com), outlines the key elements contributing to a team’s success. Oliver originally assigned the following weights to these factors: shooting (40%), turnovers (25%), rebounding (20%), and free throws (15%). While the exact weighting is often debated, the importance of these four factors is widely recognized. They are critical not only for understanding a team’s offensive efficiency but also for evaluating defensive performance (e.g., analyzing the opponent’s effective field goal percentage). The four factors are:

1.	Effective Field Goal Percentage (EFG%)
2.	Turnovers Committed per Possession
3.	Offensive Rebounding Percentage (ORP)
4.	Free Throw Rate (FTR)

### Effective Field Goal Percentage (EFG%)
Effective field goal percentage adjusts for the difference in value between two-point and three-point shots, giving three-pointers 50% more weight since they are worth 50% more points. The formula for EFG% is:

$$
EFG = \frac{fg2m + 1.5 \times fg3m}{fga}
$$
Where 

-	`fg2m`: Number of two-point field goals made
-	`fg3m`: Number of three-point field goals made
-	`fga`: Total field goals attempted

### Turnovers Committed per Possession
This metric measures how often a team turns the ball over relative to the number of possessions they have in a game. It highlights a team’s ball-handling efficiency by considering turnovers (`tov`) in relation to total possessions (`poss`).

### Offensive Rebounding Percentage
Offensive rebounding percentage evaluates a team’s ability to secure offensive rebounds compared to the total number of available offensive rebounds. The formula is:

$$
ORP = \frac{\text{Team Offensive Rebounds}}{\text{Team Offensive Rebounds + Opponent Offensive Rebounds}}
$$

### Free Throw Rate
Free throw rate measures how frequently a team attempts free throws compared to its total field goal attempts. The formula is:

$$
FTR = \frac{\text{Free Throw Attempts (fta)}}{\text{Field Goal Attempts (fga)}}
$$

<br>

