---
title: Data Storytelling Team Project - Golf
author: Byeong-Hak Choe
institute: SUNY Geneseo
date: 2025-11-15
# toc: true
# toc-depth: 2
# toc-expand: true
# toc-title: Contents
# number-sections: false
# number-depth: 3

code-fold: false

execute:
  echo: true
  eval: true
  message: false
  warning: false
  fig-width: 9
  fig-height: 6
---
```{r}
#| include: false
library(tidyverse)
library(skimr)
library(broom)
library(hrbrthemes)

theme_set(theme_ipsum() +
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = element_text(size = rel(1.5) ),
                axis.title.y = element_text(size = rel(1.5) ))
          )

# Set global options for color-blind-friendly scales
# scale_colour_discrete <- function(...) scale_colour_viridis_d(...)
scale_colour_discrete <- function(...) scale_color_colorblind(...)
scale_fill_discrete <- function(...) scale_fill_colorblind(...)
```

# Data

The following lists data frames about golf:

- `pga_tournaments`: PGA Tournament Data from 2022 season  
- `ow_golf_rankings`: Official World Golf Rankings in June 2022  
- `pga_tournaments_avg`: Average player performance statistics with driving distance and putting by tournament.


## Average PGA Performance Data from 2022 season (`pga_tournaments_avg`)

A data frame with 1,387 rows and 12 **variables**:

  - Cases include all golfers who made the cut in each of 19 PGA tournaments in 2022.
  
```{r}
pga_tournaments_avg <- read_csv(
  "http://bcdanl.github.io/data/pga_tournaments_avg.csv")
```

```{r}
#| echo: false
rmarkdown::paged_table(pga_tournaments_avg)
```

<br>

`playerName`  
Player’s full name.

`country`  
Player’s country of origin or representation.

`avgDriveDist`  
Average driving distance off the tee (typically measured in yards).


`drivePct`  
Driving accuracy percentage (for example, the percentage of drives/fairways hit from the tee).

`avgPuttsPerRound`  
Average number of putts per 18-hole round.

`onePuttPct`  
Percentage of holes where the player needed only one putt on the green.

`driveSG`  
Strokes gained from driving (off-the-tee performance relative to the field).

`puttsSG`  
Strokes gained from putting (putting performance relative to the field).

`avgScore`  
Average 18-hole score (strokes per round). Scoring average to measure success.

`Money`  
Total prize money earned over the period or tournament (typically in U.S. dollars).

`Points`  
Tour points associated with the performance in the Fedex Cup.

`tournament`  
Tournament name or identifier associated with the statistics in the row.


<br>


## PGA Tournament Data from 2022 season (`pga_tournaments`)
A data frame with 3,676 rows and 34 **Variables**:

```{r}
pga_tournaments <- read_csv("http://bcdanl.github.io/data/pga_tournaments.csv")
```

```{r}
#| echo: false
rmarkdown::paged_table(pga_tournaments)
```

<br>

`Player_initial_last`
initial of first name and complete last name of player

`tournament.id`
tournament ID

`player.id`
plyaer ID

`hole_par`
par across all holes played by the player

`strokes`
strokes taken on all holes played by the player

`score_relative_to_par` 
Score relative to par (`strokes` - `hole_par`)

`hole_DKP`
Draftkings points on holes

`hole_FDP`
Fanduel points on the holes

`hole_SDP`
Showdown points on the holes

`streak_DKP`
Draftkings points on streaks

`streak_FDP`
Fanduel points on streaks

`streak_SDP`
Showdown points on streaks

`n_rounds`
number of rounds played

`made_cut`
player made the cut or not

`pos`
finishing position of player

`finish_DKP`
Draftkings points on the finishing position

`finish_FDP`
Fanduel points on the finishing position

`finish_SDP`
Showdown points on the finishing position

`total_DKP`
total Draftkings points for the tournament

`total_FDP`
total Fanduel points for the tournament

`total_SDP`
total Showdown points for the tournament

`player`
player full name

`tournament.name`
tournament full name

`course`
course name

`date`
data of tournament

`purse`
total prize money (in millions)

`season`
year of season

`no_cut`
Not sure

`Finish`
finishing position

`sg_putt`
strokes gained from putting

`sg_arg`
strokes gained around the green

`sg_app`
strokes gained on approach shots

`sg_ott`
strokes gained off the tee

`sg_t2g`
strokes gained tee to green

`sg_total`
total strokes gained


Source: [https://datagolf.com](https://datagolf.com)




## Official World Golf Rankings in June 2022  (`ow_golf_rankings`)
- A data frame with 300 rows and 2 **Variables**:

```{r}
ow_golf_rankings <- read_csv("http://bcdanl.github.io/data/ow_golf_rankings.csv")
```

```{r}
#| echo: false
rmarkdown::paged_table(ow_golf_rankings)
```

<br>

`WGR_June_2022`
official golf world ranking

`Player_initial_last`
initial of first name and complete last name of player


Source: [https://www.owgr.com/](https://www.owgr.com/)


# Explaining Key Golf Variables 

This section explains the meaning of important variables found in `pga_tournaments` and provides context for people who do **not** know golf.  


---

## 🏌️ Basic Golf Concepts

Before interpreting the variables, here are two essential ideas:

### 1. A “stroke” is one hit of the ball.
Fewer strokes = better scoring.

### 2. Every hole has a “par.”
Par is the expected number of strokes a skilled player should take:  
- 3 strokes → par  
  - **Par**: target number of strokes for a hole (usually 3, 4, or 5).  


- `strokes == hole_par` → “made par”  
- `strokes = par – 1` → birdie (good)  
- `strokes = par + 1` → bogey (bad)


**Idea:** strokes gained numbers tell us whether the player is **better or worse than the PGA Tour average** for that type of shot.  

- Positive value → better than average  
- Negative value → worse than average

### 3. Tournaments have multiple rounds.
Most events have 4 rounds (Thu–Sun), each of 18 holes.

### 4. There is a “cut” after round 2.
Only the best players continue to round 3–4. Others “miss the cut.”

- **Tournament**: usually 4 rounds (4 × 18 holes).  
- **Cut**: after 2 rounds, only players with the best scores continue to rounds 3–4.




---

# 🟢 Variable Explanations + What We Can Analyze

## 1. `strokes`
**Meaning**:
Number of strokes the player took on a single hole.

**Golf context**:
This is the fundamental scoring unit in golf.

**Analysis examples**:

- Average strokes per round  
- Birdie/bogey rates  
- Score relative to par (`strokes - hole_par`)  
- Compare performance on par-3 vs par-4 vs par-5 holes  

---

## 2. `hole_par`
**Meaning**:
Par value of the hole

**Golf context**:
Indicates the expected difficulty of the hole.

**Analysis examples**:

- Compute score relative to par  
- Identify which par type a player scores best on  
- Hole difficulty analysis across tournaments  

---

## 3. `made_cut`
**Meaning**:
Indicator of whether the player advanced past the cut after 2 rounds.

**Golf context**:
If you make the cut, you continue playing the weekend.

**Analysis examples**:

- Probability of making the cut  
- Skill differences between cut-makers and non-cut-makers  
- Effect of cuts on fantasy scoring or prize money  

---

## 4. `n_rounds`
**Meaning**:
Number of rounds a player completed.

**Golf context**:
Usually 2 if cut, 4 if not cut (unless it’s a no-cut event).

**Analysis examples**:

- Relation between number of rounds and finishing position  
- Identify consistent 4-round players  
- Evaluate endurance or weekend performance  

---

## 5. Strokes Gained Metrics
**Variables**:  
`sg_putt`, `sg_arg`, `sg_app`, `sg_ott`, `sg_t2g`, `sg_total`

**Meaning**:
These measure how many strokes a player gained or lost compared to the PGA Tour average on specific types of shots.

**Categories**:

- **SG:OTT** — Off the tee (drives)  
- **SG:APP** — Approach shots to the green  
- **SG:ARG** — Around the green (chipping)  
- **SG:PUTT** — Putting  
- **SG:T2G** — All shots except putting  
- **SG:TOTAL** — Overall performance

**Analysis examples**:

- Identify player strengths and weaknesses  
- Radar charts to compare skills  
- Compare top-50 players vs outside top-50  
- Which skill areas drive finishing position?  

---

## 6. `pos` / `Finish`
**Meaning**:
Player’s final ranking in the tournament (1 = winner).

**Analysis examples**:

- Trends in finishing positions over time  
- Correlation with strokes gained  
- Compare finishing position across different courses  

---

## 7. Tournament Metadata
**Variables**:  
`tournament.id`, `tournament.name`, `course`, `date`, `season`

**Meaning**:
Information describing the tournament.

**Analysis examples**:

- Course difficulty comparisons  
- Seasonal patterns in scoring  
- Performance on specific courses (e.g., Augusta vs Torrey Pines)  

---

## 8. Fantasy Scoring Variables
`hole_DKP`, `hole_FDP`, `hole_SDP`  
`streak_DKP`, `streak_FDP`, `streak_SDP`  
`finish_DKP`, `finish_FDP`, `finish_SDP`  
`total_DKP`, `total_FDP`, `total_SDP`

**Meaning**:
Fantasy golf points for DraftKings (DKP), FanDuel (FDP), and SuperDraft (SDP).

**Golf context**:
Fantasy scoring rewards:
- Birdies  
- Eagles  
- Streaks  
- Finishing position  
- Penalties for bogeys

**Analysis examples**:

- Compare fantasy scoring to traditional golf scoring  
- Identify players who benefit from fantasy-scoring rules  
- Optimize fantasy lineup selection  

---

## 9. `purse`
**Meaning**:
Total prize money available in the tournament.

**Analysis examples**:

- Relationship between purse size and player performance  
- Higher-purse events vs field strength  
- Earnings distribution across seasons  

---

## 10. `no_cut`
**Meaning**:
Indicator whether the tournament has no cut.

**Golf context**:
Some events allow all players to play all rounds.

**Analysis examples**:

- Compare scoring in cut vs no-cut events  
- Effect on strokes gained and finishing positions  
- Fantasy scoring differences  

---

# Summary Table

| Variable | Plain-English Explanation | Possible Analyses |
|---------|---------------------------|-------------------|
| `strokes` | Shots taken on a hole | Scoring average, consistency, birdies/bogeys |
| `hole_par` | Difficulty benchmark | Score relative to par, hole difficulty |
| `made_cut` | Player advanced to weekend? | Cut probability, skill gap, fantasy impact |
| `n_rounds` | Number of rounds played | Consistency, performance over rounds |
| Strokes gained metrics | Skill measurement vs PGA average | Player skill profiles, radar charts |
| `pos` / `Finish` | Final tournament rank | Trend analysis, correlation with skills |
| Tournament metadata | Info on course & event | Course difficulty, seasonal trends |
| Fantasy scoring | DK/FanDuel/SuperDraft points | Lineup optimization, scoring comparisons |
| `purse` | Total prize money | Incentives, field strength |
| `no_cut` | Cut or no-cut event | Scoring pattern differences |

---

# Conclusion

Even without deep golf knowledge, these variables allow meaningful analysis of:

- Player performance  
- Tournament structure  
- Scoring behavior  
- Skill strengths and weaknesses  
- Course difficulty  
- Fantasy scoring strategies  

This dataset supports a full range of **descriptive**, **comparative**, and **strategic** analytics.