---
title: Data Storytelling Team Project - Soccer
author: Byeong-Hak Choe
institute: SUNY Geneseo
date: 2025-11-15
# toc: true
# toc-depth: 2
# toc-expand: true
# toc-title: Contents
# number-sections: false
# number-depth: 3

code-fold: false

execute:
  echo: true
  eval: true
  message: false
  warning: false
  fig-width: 9
  fig-height: 6
---
```{r}
#| include: false
library(tidyverse)
library(skimr)
library(broom)
library(hrbrthemes)

theme_set(theme_ipsum() +
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = element_text(size = rel(1.5) ),
                axis.title.y = element_text(size = rel(1.5) ))
          )

# Set global options for color-blind-friendly scales
# scale_colour_discrete <- function(...) scale_colour_viridis_d(...)
scale_colour_discrete <- function(...) scale_color_colorblind(...)
scale_fill_discrete <- function(...) scale_fill_colorblind(...)
```

# Data

The following lists data frames about English Premier League (EPL):

- `epl_teams`: team statistics for the English Premier League 2017-18 through 2024-25
- `epl_players`: Individual player statistics for the English Premier League's 2021-22 through 2024-25
- `epl_gks`: Individual goalkeeper statistics for the English Premier League's 2021-22 through 2024-25



## EPL Team Statistics 2017-18 through 2024-25 (`epl_teams`)
- `epl_teams`: team statistics for the English Premier League the English Premier League 2017-18 through 2024-25
  - A dataset containing team statistics for the English Premier League
  - A data frame with 160 rows and 29 variables

```{r}
epl_teams <- read_csv("http://bcdanl.github.io/data/epl_teams_2017-18_2024-25.csv")
```

```{r}
#| echo: false
rmarkdown::paged_table(epl_teams)
```

<br>

`squad`
Team or club name

`season_end_year`
Season end year (e,g., 2025 if the season is 2024-25)


`num_players`
Number of players used in the season

`age`
Age of players weighted by minutes played

`poss`
Possession as a percentage of passes attempted

`goals`
Goals scored by team

`assists`
Assists by team

`goals_assists`
Goals scored plus assists by team

`non_pen_goals`
Non-penalty goals scored by team

`pen_goals`
Penalty kick goals scored by team

`pen_att`
Penalty kick attempts by team

`yellow_cards`
Yellow cards earned by team

`red_cards`
Red cards earned by team

`xG`
Expected goals

`non_pen_xG`
Non-penalty kick expected goals

`xA`
Expected assists

`non_pen_xGxA`
Non-penalty kick expected goals and assists

`prg_carry`
Progressive carries

`prg_pass`
Progressive passes

`goals_p90`
Goals per ninety minutes

`assists_p90`
Assists per ninety minutes

`goals_assists_p90`
Goals and assists per ninety minutes

`non_pen_goals_p90`
Non-penalty kick goals per ninety minutes

`non_pen_goals_assists_p90`
Non-penalty kick goals and assists per ninety minutes

`xG_p90`
Expected goals per ninety minutes

`xA_p90`
Expected assists per ninety minutes

`xG_xA_p90`
Expected goals plus assists per ninety minutes

`non_pen_xG_p90`
Expected goals minus penalty goals per ninety minutes

`non_pen_xG_xA_p90`
Expected goals plus assists minus penalty goals per ninety minutes

Source: [https://fbref.com/en/comps/9/2021-2022/stats/2021-2022-Premier-League-Stats](https://fbref.com/en/comps/9/2021-2022/stats/2021-2022-Premier-League-Stats)




## EPL Player Statistics 2021-22 through 2024-25  (`epl_players`)

- `epl_players`: Individual player statistics for the English Premier League's 2021-22 through 2024-25
  - A data frame with 2,268 observations on the 38 variables.

```{r}
epl_players <- read_csv("http://bcdanl.github.io/data/epl_players_all_w_salary_2021-22_2024-25.csv")
```

```{r}
#| echo: false
rmarkdown::paged_table(epl_players)
```

<br>


`season_end_year`
Season end year (e,g., 2025 if the season is 2024-25)

`player`
Player name

`nation`
Player home country

`pos`
Player position

`squad`
Player team

`gross_salary_weekly_pound`
Weekly salary in pound (£)

`gross_salary_yearly_pound`
Yearly salary in pound (£)


`age`
Age of player

`born`
Birth year of player

`mp`
Matches played

`starts`
Number of matches in which player started the game

`min`
Total minutes played in the season

`min_p90`
Total minutes played in the season per ninety minutes

`goals`
Goals scored by player

`assists`
Assists by player

`goals_assists`
Goals scored plus assists by player

`non_pen_goals`
Non-penalty goals scored by player

`pen_goals`
Penalty kick goals scored by player

`pen_att`
Penalty kick attempts by player

`yellow_cards`
Yellow cards earned by player

`red_cards`
Red cards earned by player

`xG`
Expected goals

`non_pen_xG`
Non-penalty kick expected goals

`xA`
Expected assists

`non_pen_xGxA`
Non-penalty kick expected goals and assists

`prg_carry`
Progressive carries

`prg_pass`
Progressive passes

`prg_reception`
Progressive passes received

`goals_p90`
Goals per ninety minutes

`assists_p90`
Assists per ninety minutes

`goals_assists_p90`
Goals and assists per ninety minutes

`non_pen_goals_p90`
Non-penalty kick goals per ninety minutes

`non_pen_goals_assists_p90`
Non-penalty kick goals and assists per ninety minutes

`xG_p90`
Expected goals per ninety minutes

`xA_p90`
Expected assists per ninety minutes

`xG_xA_p90`
Expected goals plus assists per ninety minutes

`non_pen_xG_p90`
Expected goals minus penalty goals per ninety minutes

`non_pen_xG_xA_p90`
Expected goals plus assists minus penalty goals per ninety minutes

Source: [https://fbref.com/en/comps/9/2021-2022/stats/2021-2022-Premier-League-Stats](https://fbref.com/en/comps/9/2021-2022/stats/2021-2022-Premier-League-Stats)


##  EPL Goalkeeper Statistics 2021-22 through 2024-25 (`epl_gks`)

- `epl_gks`: Individual goalkeeper statistics for the English Premier League's 2021-22 through 2024-25
  - A data frame with 165 rows and 33 variables:

```{r}
epl_gks <- read_csv("http://bcdanl.github.io/data/epl_gk_stats_2021-22_2024-25.csv")
```

```{r}
#| echo: false
rmarkdown::paged_table(epl_gks)
```

<br>

`season_end_year`
Season end year (e,g., 2025 if the season is 2024-25)

`player`
Player name

`nation`
Player home country

`pos`
Player position

`squad`
Player team

`age`
Age of player

`born`
Birth year of player

`min_p90`
Total minutes played in the season per ninety minutes

`goals`
Goals scored against player

`pk_allowed`
Penalty kick goals scored against player

`free_kick_goals`
Free kick goals scored against player

`corner_kick_goals`
Corner kick goals scored against player

`own_goals_against`
Own goals scored against player

`post_shot_xG`
Post-shot expected goals

`post_shot_xG_per_shot_on_target`
Post-shot expected goals per shot on target

`post_shot_xG_minus_goals`
Post-shot expected goals minus goals per shot on target

`post_shot_xG_minus_goals_p90`
Post-shot expected goals minus goals per shot on target per ninety minutes

`launch_completed`
Passes longer than 40 yards (launch) completed

`launch_attempt`
Passes longer than 40 yards attempted

`launch_completion_percent`
Passes longer than 40 yards (launch) completion percentage

`pass_attempt`
Passes attempted

`throw_attempt`
Throws attempted

`pass_launch_percent`
Percentage of passes launched

`avg_pass_len`
Average pass length

`goal_kick_attempt`
Goal kick attempts

`gk_launch_percent`
Percentage of goals kicks launched

`avg_gk_len`
Average goal kick length

`crosses_faced`
Number of crosses faced

`crosses_stopped`
Number of crosses stopped

`stop_percent`
Cross stop percentage

`def_actions_out_pen_area`
Defensive actions outside the penalty area

`def_actions_out_pen_area_p90`
Defensive actions outside the penalty area per ninety minutes

`avg_dist_def_actions`
Average distance from goal of defensive actions

Source: [https://fbref.com/en/comps/9/2021-2022/keepersadv/2021-2022-Premier-League-Stats](https://fbref.com/en/comps/9/2021-2022/keepersadv/2021-2022-Premier-League-Stats)



# Soccer Metrics

## Expected Goals and Assists

Soccer is renowned for its low-scoring nature. During the 2022 World Cup in Qatar, 172 goals were scored across 64 matches, averaging 2.68 goals per match—the highest average in World Cup history. Despite this, seven matches ended scoreless. The element of luck significantly influences goal scoring, and occasionally, players inadvertently score own goals. To better quantify scoring opportunities and player performance, the soccer analytics community has introduced **expected goals (xG)** as a metric to measure the likelihood of a shot resulting in a goal.

Various models calculate xG, but fundamentally, it represents the probability of scoring from a specific spot on the field when a shot is taken. Factors considered in xG calculations include:

- **Body Part Used**: Whether the shot was taken with the head or foot.
- **Type of Play**: How the player received the ball (e.g., cross, set piece).
- **Goalkeeper Positioning**: The location and readiness of the goalkeeper.
- **Player Positioning**: Positions of teammates and opponents between the shooter and the goal.
- **Defensive Pressure**: How tightly defenders are marking the shooter.
- **Additional Situational Variables**: Other measurable factors that may affect shot success.

Since most shots do not result in goals, and many passes to shooters lead to missed opportunities, an analogous metric called **expected assists (xA)** evaluates the likelihood that a pass will lead to a goal. These metrics are often normalized per 90 minutes—denoted as xG per 90 or xA per 90—to allow fair comparisons between players with varying playing times.

Players consistently achieving high xG values are those who frequently position themselves advantageously on the field. Coaches and scouts use xG and xA metrics to assess whether a player's performance is influenced by luck over multiple games and to evaluate offensive skills beyond basic statistics.


## Soccer Metrics

### Offensive Metrics

| Statistic                                  | Definition                                                                           |
|--------------------------------------------|--------------------------------------------------------------------------------------|
| **Goals**                                  | Total number of goals scored by the player/team                                      |
| **Assists**                                | Total number of assists made by the player/team                                      |
| **Goals and Assists**                      | Sum of goals and assists                                                             |
| **Non-Penalty Goals**                      | Goals scored excluding penalty kicks                                                 |
| **Penalty Goals**                          | Goals scored from penalty kicks                                                      |
| **Penalty Attempts**                       | Number of penalty kick attempts                                                      |
| **Expected Goals (xG)**                    | Expected number of goals based on shot quality and other factors                     |
| **Non-Penalty xG**                         | Expected goals excluding penalty kicks                                               |
| **Expected Assists (xA)**                  | Expected number of assists based on pass quality and other factors                   |
| **Non-Penalty xG + xA**                    | Sum of non-penalty expected goals and expected assists                               |
| **Progressive Carries**                    | Times a player significantly moves the ball towards the opponent's goal              |
| **Progressive Passes**                     | Forward passes advancing the ball significantly towards the opponent's goal          |
| **Goals per 90 Minutes**                   | Average number of goals scored per 90 minutes played                                 |
| **Assists per 90 Minutes**                 | Average number of assists per 90 minutes played                                      |
| **Goals and Assists per 90 Minutes**       | Sum of goals and assists per 90 minutes played                                       |
| **Non-Penalty Goals per 90 Minutes**       | Non-penalty goals scored per 90 minutes played                                       |
| **Non-Penalty Goals and Assists per 90**   | Sum of non-penalty goals and assists per 90 minutes played                           |
| **xG per 90 Minutes**                      | Expected goals per 90 minutes played                                                 |
| **xA per 90 Minutes**                      | Expected assists per 90 minutes played                                               |
| **xG + xA per 90 Minutes**                 | Sum of expected goals and assists per 90 minutes played                              |
| **Non-Penalty xG per 90 Minutes**          | Non-penalty expected goals per 90 minutes played                                     |
| **Non-Penalty xG + xA per 90 Minutes**     | Sum of non-penalty expected goals and assists per 90 minutes played                  |

### Defensive Metrics

| Statistic                                  | Definition                                                                           |
|--------------------------------------------|--------------------------------------------------------------------------------------|
| **Yellow Cards**                           | Number of yellow cards received                                                      |
| **Red Cards**                              | Number of red cards received                                                         |
| **Progressive Passes Received**            | Passes received that advance the ball significantly towards the opponent's goal      |

### Goalkeeping Metrics

| Statistic                                  | Definition                                                                           |
|--------------------------------------------|--------------------------------------------------------------------------------------|
| **Goals Against**                          | Total number of goals conceded                                                       |
| **Penalty Kicks Allowed**                  | Goals conceded from penalty kicks                                                    |
| **Free Kick Goals Against**                | Goals conceded from free kicks                                                       |
| **Corner Kick Goals Against**              | Goals conceded from corner kicks                                                     |
| **Own Goals Against**                      | Goals scored by teammates into their own net                                         |
| **Post-Shot xG**                           | Expected goals based on the quality of shots on target faced                         |
| **Post-Shot xG per Shot on Target**        | Average post-shot expected goals per shot on target faced                            |
| **Post-Shot xG Minus Goals**               | Difference between post-shot expected goals and actual goals conceded                |
| **Post-Shot xG Minus Goals per 90 Minutes**| Post-shot xG minus goals conceded per 90 minutes                                     |
| **Crosses Faced**                          | Number of crosses faced by the goalkeeper                                            |
| **Crosses Stopped**                        | Number of crosses successfully stopped by the goalkeeper                             |
| **Cross Stop Percentage**                  | Percentage of crosses stopped (Crosses Stopped ÷ Crosses Faced)                      |
| **Defensive Actions Outside Penalty Area** | Actions by the goalkeeper outside the penalty area                                   |
| **Defensive Actions Outside PA per 90**    | Such actions per 90 minutes                                                          |
| **Average Distance of Defensive Actions**  | Average distance from goal of the goalkeeper's defensive actions                     |

### Team Metrics

| Statistic                                  | Definition                                                                           |
|--------------------------------------------|--------------------------------------------------------------------------------------|
| **Squad**                                  | Team or club name                                                                    |
| **Season**                                 | Season                                                                               |
| **Number of Players**                      | Number of players used in the season                                                 |
| **Average Age**                            | Age of players weighted by minutes played                                            |
| **Possession Percentage**                  | Percentage of possession based on passes attempted                                   |
