---
title: Data Storytelling Team Project - Football
author: Byeong-Hak Choe
institute: SUNY Geneseo
date: last-modified

toc: true
toc-depth: 2
toc-expand: true
toc-title: Contents
number-sections: false
number-depth: 3

code-fold: false

execute:
  echo: true
  eval: true
  message: false
  warning: false
  fig-width: 9
  fig-height: 6
---
```{r}
#| include: false
library(tidyverse)
library(skimr)
library(broom)
library(hrbrthemes)

theme_set(theme_ipsum() +
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = element_text(size = rel(1.5) ),
                axis.title.y = element_text(size = rel(1.5) ))
          )

# Set global options for color-blind-friendly scales
# scale_colour_discrete <- function(...) scale_colour_viridis_d(...)
scale_colour_discrete <- function(...) scale_color_colorblind(...)
scale_fill_discrete <- function(...) scale_fill_colorblind(...)
```



# Data

The following lists data frames about National Football League (NFL) for the seasons from 2014-15 through 2023-2024:

- `nfl_team_epa`: Team's mean expected points added (EPA) when the team was on offense and when the team was on defense
  - For the details about EPA, please refer to the Football Metrics section below in the webpage.
- `nfl_field_goals`: Play-by-play statistics at situations when field goals were attempted during the game
- `nfl_passers`: weakly EPA and completion percentage over expected (CPOE) among players who passed more than 44 times within a week
- `nfl_player_stat`: Player statistics
- `nfl_receivers`:  Total EPA of players whose positions are either "WR", "TE", or "RB", and top 10 players in terms of total EPA

| Position | Full Name        | Main Role                                 | Skills Required                   |
|----------|------------------|------------------------------------------|-----------------------------------|
| **WR**   | Wide Receiver    | Catch passes and gain yards or score TDs | Speed, agility, reliable hands    |
| **TE**   | Tight End        | Block defenders and catch passes         | Strength, versatility, reliable hands |
| **RB**   | Running Back     | Run the ball, catch passes, block        | Speed, vision, agility            |
| **QB**   | Quarterback      | Lead the offense and throw passes        | Decision-making, accuracy, arm strength, leadership |




## NFL Team EPA (`nfl_team_epa`)
- `nfl_team_epa`: Team's mean EPA when the team was on offense and when the team was on defense

```{r}
nfl_team_epa <- read_csv("http://bcdanl.github.io/data/nfl_team_epa.csv")
```

```{r}
#| echo: false
rmarkdown::paged_table(nfl_team_epa)
```

<br>

`season`
Starting year of the season (2014 if 2014-15 season)

`team`
Team abbreviation

`off_epa`
Offensive EPA

`def_epa`
Defensive EPA


## NFL Field Goals  (`nfl_field_goals`)

- `nfl_field_goals`: Play-by-play statistics at situations when field goals were attempted during the game
  - A data frame with 10047 observations on the 23 variables.

```{r}
nfl_field_goals <- read_csv("http://bcdanl.github.io/data/nfl_field_goals.csv")
```

```{r}
#| echo: false
rmarkdown::paged_table(nfl_field_goals)
```

<br>


```{r}
#| echo: false
DT::datatable(read_csv("https://bcdanl.github.io/data/nfl_field_goals_vars.csv"),
  options = list(
  pageLength = 25))
```

<br>

- `fg_distance` accurately reflects the total distance of a field goal attempt:
  - The total addition of 17 yards comes from 10 yards (end zone) + 7 yards (holder’s position).

<br>


##  NFL Passer's EPA and COPE (`nfl_passers`)

- `nfl_passers`: weakly mean EPA and completion percentage over CPOE among players who passed more than 44 times within a week
  - A data frame with 1098 rows and 22 variables:

```{r}
nfl_passers <- read_csv("http://bcdanl.github.io/data/nfl_passers.csv")
```

```{r}
#| echo: false
rmarkdown::paged_table(nfl_passers)
```

<br>

## NFL Player Statistics (`nfl_players_stat`)
- `nfl_players_stat`: Player statistics

```{r}
nfl_players_stat <- read_csv("http://bcdanl.github.io/data/nfl_players_stat.csv")
```

```{r}
#| echo: false
rmarkdown::paged_table(nfl_players_stat)
```

<br>

`yards` `rushing_yards` + `receiving_yards`

`rushing_yards` Yards gained when rushing with the ball (incl. scrambles and kneel downs). Also includes yards gained after obtaining a lateral on a play that started with a rushing attempt.

`receiving_yards` Yards gained after a pass reception. Includes yards gained after receiving a lateral on a play that started as a pass play.

`touches` `carries` + `receptions`

`carries` The number of official rush attempts (incl. scrambles and kneel downs). Rushes after a lateral reception don't count as carry.

`receptions` The number of pass receptions. Lateral receptions officially don't count as reception.

`tds` `rushing_tds` + `receiving_tds`

`rushing_tds` The number of rushing touchdowns (incl. scrambles). Also includes touchdowns after obtaining a lateral on a play that started with a rushing attempt.

`receiving_tds` The number of touchdowns following a pass reception. Also includes touchdowns after receiving a lateral on a play that started as a pass play.


## NFL Receivers (`nfl_receivers`)

- `nfl_receivers`:  Total EPA of players whose positions are either "WR", "TE", or "RB", and top 10 players in terms of total EPA

```{r}
nfl_receivers <- read_csv("http://bcdanl.github.io/data/nfl_receivers.csv")
```

```{r}
#| echo: false
rmarkdown::paged_table(nfl_receivers)
```

<br>

`epa_rank`: Ranking in terms of `tot_epa` (The lower `tot_epa`, the higher EPA)

`epa_rank_within_position`: Ranking in terms of EPA within the group of the same position

`n_received`: the number of passes a player received

`tot_epa`: Total EPA within a season

<br>

# Football Metrics


## Expected Points

Expected Points Added (EPA) is a football analytics metric that measures the value of a play in terms of its impact on the team’s expected scoring. It quantifies how much a single play increases or decreases a team’s chances of scoring compared to the situation before the play.

### How EPA Works

Every play in football occurs within a specific context (down, distance, field position, time remaining, and score). Historical data is used to calculate the **expected points (EP)** a team can expect to score from their current situation. **EPA** is the difference between the expected points after the play and before the play.

### Formula:

$$
EPA = EP (\text{after the play}) - EP (\text{before the play})
$$

### Key Insights:

1. Positive EPA: The play improved the team’s scoring chances.
  - Example: A 20-yard pass on 3rd and 8 increases the likelihood of scoring.
2. Negative EPA: The play reduced the team’s scoring chances.
	-	Example: A sack or an interception harms the team’s scoring potential.

### Why EPA Is Important

-Contextual: Accounts for situational factors, making it more informative than raw stats like yards gained.
-Play Evaluation: Helps determine the effectiveness of specific plays or players.
-Strategic Decisions: Assists coaches and analysts in evaluating decisions like when to go for it on 4th down.

### Applications

-Offensive EPA: Evaluates how well a team’s offense increases scoring opportunities.
-Defensive EPA: Measures how effectively a defense reduces the opposing team’s scoring potential.
-Player Performance: Used to assess quarterbacks, running backs, wide receivers, and defenders by their contribution to scoring or preventing points.


## Completion Percentage Over Expected (CPOE)

Completion Percentage Over Expected (CPOE) is a football analytics metric that evaluates a quarterback’s passing performance by comparing their actual completion percentage to the expected completion percentage based on the difficulty of their pass attempts.

### How CPOE Works

1.	Actual Completion Percentage (COMP%):
	-	The percentage of passes a quarterback completes.
2.	Expected Completion Percentage (xCOMP%):
	-	Calculated based on factors like:
	-	Distance of the throw (air yards)
	-	Angle of the throw
	-	Receiver separation
	-	Defensive pressure
	-	Game situation (e.g., down, distance, and field position)
	-	Derived from historical data on similar passes.
3.	CPOE Formula:
$$
CPOE = COMP - xCOMP
$$


### Interpretation of CPOE
-	Positive CPOE: Indicates the quarterback is completing more passes than expected, showcasing accuracy and skill.
-	Negative CPOE: Indicates the quarterback is completing fewer passes than expected, potentially highlighting issues with accuracy or decision-making.

### Why CPOE Is Useful

-	Isolates Skill: It accounts for the difficulty of throws, focusing on the quarterback’s performance rather than the system or play design.
-	Complementary Metric: Often paired with EPA/play to provide a comprehensive evaluation of a quarterback’s impact.
-	Game Context: Helps differentiate between quarterbacks who excel in challenging situations versus those whose stats are inflated by easy throws.
