---
title: Homework Assignment 5
subtitle: ""
date: last-modified
from: markdown+emoji
comments: false
code-fold: false

execute: 
  warning: false
  message: false
  fig-width: 9
  fig-height: 7
  fig-align: center
  eval: false
---


```{r}
#| include: false
#| eval: true

library(tidyverse)
library(skimr)
library(DT)
library(hrbrthemes)
library(ggthemes)
library(gapminder)

titanic <- read_csv("https://bcdanl.github.io/data/titanic_cleaned.csv")
oj <- read_csv("https://bcdanl.github.io/data/dominick_oj_na.csv")
mlb_bat <- read_csv("https://bcdanl.github.io/data/MLB_batting.csv")

theme_set(theme_ipsum() +
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = element_text(size = rel(1.5)),
                axis.title.y = element_text(size = rel(1.5), angle = 0),
                legend.title = element_text(size=rel(1.25))
                ))

# Set global options for color-blind-friendly scales
scale_colour_discrete <- function(...) scale_color_tableau(...)
scale_fill_discrete <- function(...) scale_fill_tableau(...)

```


- Please answer all of the following questions thoroughly, ensuring that none are left unanswered.
- **Craft your answers in a Word document and submit your Word document to Brightspace**
- For Homework Assignment 5, the use of any generative artificial intelligence (AI) tools is strictly required.
  - Pick one generative AI tool you prefer.
  - Copy your conversation with generative AI and paste it to your Word document.
- You may submit multiple times, but only your most recent submission will be evaluated.
- The due is <u>December 10, 11:59 P.M., 2024, Tuesday, Eastern Time</u>.

<br>

## Question 1.
What generative AI tool have you used for this homework assignment?


<br>

## Question 2.
- Provide your conversation with generative AI to do the following tasks:
  - Translate the following R `ggplot` code into Python `seaborn` code to generate a scatter plot showing the relationship between “sales” and “price” using the CSV file, `http://bcdanl.github.io/data/dominick_oj_feat.csv`.
  - Make a step-by-step comparison between the Python code and the R code to understand how each part corresponds to the other.


```{r}
#| echo: true
#| eval: false

library(tidyverse)
oj <- read_csv("http://bcdanl.github.io/data/dominick_oj_feat.csv")
ggplot(data = oj, 
       mapping = aes(x = price, y = sales,
                     color = brand)) +
  geom_point(alpha = .3) + 
  geom_smooth(method = "lm") +
  labs(title = "Scatter Plot of Sales vs. Price",
       x = "Price",
       y = "Sales")
```

<br>

## Question 3.
- Provide your conversation with generative AI to do the following task:
  - Make an in-line comment on each line of the following R `ggplot` code.


```{r}
library(tidyverse)
oj <- read_csv("http://bcdanl.github.io/data/dominick_oj_feat.csv")
ggplot(data = oj, 
       mapping = aes(x = price, y = sales,
                     color = brand)) +
  geom_point(alpha = .3) + 
  geom_smooth(method = "lm") +
  labs(title = "Scatter Plot of Sales vs. Price",
       x = "Price",
       y = "Sales")
```

<br>

## Question 4
Insert an image file of your scatterplot showing `sales` versus `price`, with data points color-coded by `brand`, generated using Power BI.

<br>





## Question 5.
- Provide your conversation with generative AI to debug the following code:
  - Explain to the generative AI the error message you received when running the code below:

```{r}
#| eval: false

oj |> 
  counting(brand, ad_status)
```



<br>

## Question 6.
-	Provide your conversation with generative AI for adding a new variable, `revenue`, to the `oj` data frame.
    - The `revenue` variable should be computed as the product of `sales` and `price` to provide information about the weekly revenue for each orange juice brand.

```{r}
library(tidyverse)
oj <- read_csv("http://bcdanl.github.io/data/dominick_oj_feat.csv")
```



<br>
