---
title: Homework 4
subtitle: Data Transformation & Visualization in R — Part II
date: 2025-11-20
from: markdown+emoji
comments: false
code-fold: false

execute: 
  warning: false
  message: false
  fig-width: 9
  fig-height: 7
  fig-align: center
  eval: false
  
toc: true
---

```{r}
#| include: false
#| eval: true

library(tidyverse)
library(skimr)
library(DT)
library(hrbrthemes)
library(ggthemes)
library(gapminder)

titanic <- read_csv("https://bcdanl.github.io/data/titanic_cleaned.csv")
oj <- read_csv("https://bcdanl.github.io/data/dominick_oj_na.csv")
mlb_bat <- read_csv("https://bcdanl.github.io/data/MLB_batting.csv")

theme_set(theme_ipsum() +
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = element_text(size = rel(1.5)),
                axis.title.y = element_text(size = rel(1.5), angle = 0),
                legend.title = element_text(size=rel(1.25))
                ))

# Set global options for color-blind-friendly scales
scale_colour_discrete <- function(...) scale_color_tableau(...)
scale_fill_discrete <- function(...) scale_fill_tableau(...)

```


# Multiple Choice Questions

## Question 1
When visualizing grouped data, which of the following best describes the difference between using the `aes(color = ...)` aesthetic and using facet_wrap(...)`?

```{=html}
      <div class="control">
        <label class="radio">
          <input type="radio" name="q1" value="a" required />
          <code>color</code> aesthetic mapping separates groups into different plots, while faceting overlays groups in one plot
        </label>
        <label class="radio">
          <input type="radio" name="q1" value="b" />
          <code>color</code> aesthetic mapping differentiates groups <em>within the same plot</em>, while faceting displays each group in its <em>own panel</em>
        </label>
        <label class="radio">
          <input type="radio" name="q1" value="c" />
          Both <code>color</code> aesthetic mapping and faceting always produce identical visualizations
        </label>
        <label class="radio">
          <input type="radio" name="q1" value="d" />
          Faceting is only used for time-series data, while <code>color</code> aesthetic mapping is only used for categorical data
        </label>
      </div>

```


::: {.callout-tip collapse="true"}
### Show answer

**Answer:** `color` aesthetic mapping differentiates groups _within the same plot_, while faceting displays each group in its _own panel_


**Explanation:**

-	`aes(color = ...)` keeps all groups in the same plot, distinguishing them by different colors.
-	`facet_wrap()` separates each group into its own subplot, giving each its own panel and axes.
-	Faceting is helpful when groups overlap too much in one plot or when per-group trends need to be seen clearly.


:::

<br>

## Question 2
In `ggplot2`, what does setting `scales = "free_y"` in `facet_wrap()` do?

```{=html}

      <div class="control">
        <label class="radio">
          <input type="radio" name="q2" value="a" required />
          Allows both the x and y axes to vary independently in each facet
        </label>
        <label class="radio">
          <input type="radio" name="q2" value="b" />
          Allows the <strong>y-axis</strong> scale to vary by facet while the x-axis is shared across facets
        </label>
        <label class="radio">
          <input type="radio" name="q2" value="c" />
          Allows the <strong>x-axis</strong> scale to vary by facet while the y-axis is shared across facets
        </label>
        <label class="radio">
          <input type="radio" name="q2" value="d" />
          Keeps both axes fixed to the same scale across all facets (the default)
        </label>
      </div>

```



::: {.callout-tip collapse="true"}
### Show answer

**Answer:** Allows the **y-axis** scale to vary by facet while the x-axis is shared across facets

**Explanation:**

-	`scales = "free_y"` means each facet gets its own y-axis scale, but all facets share the same x-axis.
-	This is useful when groups differ greatly in magnitude on the y-axis, preventing small groups from appearing flat.

:::
<br>

## Question 3
Which of the following statements about data visualizations is true?

```{=html}
<div class="control">
  <label class="radio" style="display: block;">
    <input type="radio" name="q3" value="a" /> They always explain the underlying reasons behind data trends.
  </label>
  <label class="radio" style="display: block;">
    <input type="radio" name="q3" value="b" /> They are sufficient on their own to drive data-informed decisions.
  </label>
  <label class="radio" style="display: block;">
    <input type="radio" name="q3" value="c" /> They are useful for showing “what” is happening in the data.
  </label>
  <label class="radio" style="display: block;">
    <input type="radio" name="q3" value="d" /> They eliminate the need for narratives in data storytelling.
  </label>
</div>

```



::: {.callout-tip collapse="true"}
### Show answer 
**Answer:** They are useful for showing “what” is happening in the data.

**Explanation:**
Data visualizations are excellent tools for understanding “what” is happening, but they often do not explain “why” trends occur, which requires additional analysis or narratives.
:::

<br>

## Question 4
In the ggplot2 syntax `ggplot(data, aes(x, y)) + geom_point()`, what does `aes()` stand for?

```{=html}
<div class="control">
  <label class="radio" style="display: block;">
    <input type="radio" name="q4" value="a" /> Aesthetic mappings
  </label>
  <label class="radio" style="display: block;">
    <input type="radio" name="q4" value="b" /> Arithmetic expressions
  </label>
  <label class="radio" style="display: block;">
    <input type="radio" name="q4" value="c" /> Axis equal scaling
  </label>
  <label class="radio" style="display: block;">
    <input type="radio" name="q4" value="d" /> Average error squares
  </label>
</div>

```



::: {.callout-tip collapse="true"}
### Show answer
**Answer:** Aesthetic mappings

**Explanation:**
The `aes()` function in **ggplot2** stands for aesthetic mappings, linking data variables to visual properties like `x`, `y`, `color`, `fill` or `shape`.

:::

<br>

## Question 5
The equation $\Delta\log(x) \approx \Delta x / x_{0}$ demonstrates that a small change in the natural logarithm of $x$, from an initial value $\log(x_0)$ to an ending value $\log(x_{1})$, where the change in $x$ is given by $\Delta x = x_{1} - x_{0}$, can be approximated by:

```{=html}
<div class="control">
  <label class="radio" style="display: block;">
    <input type="radio" name="q5" value="a" /> The absolute change in <span class="math inline">\(x\)</span>
  </label>
  <label class="radio" style="display: block;">
    <input type="radio" name="q5" value="b" /> The percentage change in <span class="math inline">\(x\)</span>
  </label>
  <label class="radio" style="display: block;">
    <input type="radio" name="q5" value="c" /> The square of <span class="math inline">\(x\)</span>
  </label>
  <label class="radio" style="display: block;">
    <input type="radio" name="q5" value="d" /> The reciprocal of <span class="math inline">\(x\)</span>
  </label>
</div>

```



::: {.callout-tip collapse="true"}
### Show answer
**Answer:** The percentage change in $x$

**Explanation:**
When $x$ changes by a small amount, the natural logarithm of $x$ changes approximately by the percentage change in $x$.

:::

<br>

## Question 6
Which of the following statements about mapping and setting aesthetics in `ggplot2` is FALSE?



```{=html}
<div class="control">
  <label class="radio" style="display: block;">
    <input type="radio" name="q6" value="a" /> Mapping aesthetics involves linking data variables to visual properties within <code>aes()</code>
  </label>
  <label class="radio" style="display: block;">
    <input type="radio" name="q6" value="b" /> Setting aesthetics manually is done outside of <code>aes()</code> to set fixed visual properties.
  </label>
  <label class="radio" style="display: block;">
    <input type="radio" name="q6" value="c" /> You can set aesthetics manually within <code>aes()</code> by assigning fixed values.
  </label>
  <label class="radio" style="display: block;">
    <input type="radio" name="q6" value="d" /> Mapping aesthetics allows different categories to be represented by different colors or shapes.
  </label>
</div>

```

::: {.callout-tip collapse="true"}
### Show answer
**Answer:** You can set aesthetics manually within `aes()` by assigning fixed values.

**Explanation:**
Manually setting fixed aesthetics (e.g., `color = "red"`) should be done outside of `aes()`. Inside `aes()`, values are mapped a variable in a data.frame.

:::

<br>

## Question 7
Which of the following is NOT a way to explicitly inform ggplot about the grouping structure in a line plot?


```{=html}
<div class="control">
  <label class="radio" style="display: block;">
    <input type="radio" name="q7" value="a" /> Using the <code>group</code> aesthetic.
  </label>
  <label class="radio" style="display: block;">
    <input type="radio" name="q7" value="b" /> Using the <code>color</code> aesthetic.
  </label>
  <label class="radio" style="display: block;">
    <input type="radio" name="q7" value="c" /> Using the <code>linetype</code> aesthetic.
  </label>
  <label class="radio" style="display: block;">
    <input type="radio" name="q7" value="d" /> Using the <code>size</code> aesthetic.
  </label>
</div>

```

::: {.callout-tip collapse="true"}
### Show answer
**Answer:** Using the `size` aesthetic.

**Explanation:**
The `size` aesthetic does not explicitly inform **ggplot** about grouping in a line plot, whereas `group`, `color`, and `linetype` do.

:::
<br>

## Question 8
If you have a data frame with a date variable and want to plot a time series for each category in a variable called `group_var`, what is the minimal aesthetic mapping required in `ggplot2` to correctly plot separate lines for each group?


```{=html}
<div class="control">
  <label class="radio" style="display: block;">
    <input type="radio" name="q8" value="a" /> <code>mapping = aes(x = date_var, y = value_var)</code>
  </label>
  <label class="radio" style="display: block;">
    <input type="radio" name="q8" value="b" /> <code>mapping = aes(x = date_var, y = value_var, color = group_var)</code>
  </label>
  <label class="radio" style="display: block;">
    <input type="radio" name="q8" value="c" /> <code>mapping = aes(x = date_var, y = value_var, group = group_var)</code>
  </label>
  <label class="radio" style="display: block;">
    <input type="radio" name="q8" value="d" /> Both <code>mapping = aes(x = date_var, y = value_var, color = group_var)</code> and <code>mapping = aes(x = date_var, y = value_var, group = group_var)</code>
  </label>
</div>

```

::: {.callout-tip collapse="true"}
### Show answer
**Answer:** Both `mapping = aes(x = date_var, y = value_var, color = group_var)` and `mapping = aes(x = date_var, y = value_var, group = group_var)`

**Explanation:**
Both options correctly separate lines for each group. `group` ensures proper grouping, while `color` differentiates groups visually.

:::

<br>

## Question 9
When creating a vertical boxplot in `ggplot2`, which of the following mappings is correct?


```{=html}
<div class="control">
  <label class="radio" style="display: block;">
    <input type="radio" name="q9" value="a" /> Map both the <code>x</code> and <code>y</code> aesthetics to categorical variables.
  </label>
  <label class="radio" style="display: block;">
    <input type="radio" name="q9" value="b" /> Map a numeric variable to <code>x</code> and a categorical variable to <code>y</code>.
  </label>
  <label class="radio" style="display: block;">
    <input type="radio" name="q9" value="c" /> Map a categorical variable to <code>x</code> and a numeric variable to <code>y</code>.
  </label>
  <label class="radio" style="display: block;">
    <input type="radio" name="q9" value="d" /> Map both the <code>x</code> and <code>y</code> aesthetics to numeric variables.
  </label>
</div>

```

::: {.callout-tip collapse="true"}
### Show answer
**Answer:** Map a categorical variable to `x` and a numeric variable to `y`.

**Explanation:**
In a vertical boxplot, the x-axis typically represents categories, while the y-axis represents the numeric variable whose distribution is being summarized.

:::
<br>

## Question 10
Which of the following functions is used to apply a color-blind friendly palette to the `fill` aesthetic in `ggplot2`?


```{=html}
<div class="control">
  <label class="radio" style="display: block;">
    <input type="radio" name="q10" value="a" /> <code>scale_fill_manual()</code>
  </label>
  <label class="radio" style="display: block;">
    <input type="radio" name="q10" value="b" /> <code>scale_fill_tableau()</code>
  </label>
  <label class="radio" style="display: block;">
    <input type="radio" name="q10" value="c" /> <code>scale_color_gradient()</code>
  </label>
  <label class="radio" style="display: block;">
    <input type="radio" name="q10" value="d" /> <code>scale_x_continuous()</code>
  </label>
</div>

```

::: {.callout-tip collapse="true"}
### Show answer
**Answer:** `scale_fill_tableau()` or `scale_fill_manual()` 

**Explanation:**
- The `scale_fill_tableau()` function, part of the `ggthemes` package, extends **ggplot2** with **Tableau**-inspired themes and scales, including color-blind friendly palettes.

- The `scale_fill_manual()` function can used to apply a custom color palette, including color-blind friendly palettes, to the fill aesthetic in **ggplot2**.

:::

<br><br>

# Data Visualization with `ggplot`
The followings are the R packages for this homework assignment:

```{r}
library(tidyverse)
library(skimr)
library(ggthemes)
library(gapminder)
```


## Questions 11-17
Consider the following `titanic` data.frame for Questions 11-17:

```{r}
titanic <- read_csv("https://bcdanl.github.io/data/titanic_cleaned.csv")
```

```{r}
#| echo: false
#| eval: true
rmarkdown::paged_table(titanic)
```

<br>

### Question 11

How would you create the following data.frame, `titanic_class_survival`?

```{r}
#| echo: false
#| eval: true
#| results: asis

titanic_class_survival <- titanic |> count(class, survived)
rmarkdown::paged_table(titanic_class_survival)
```

- The `titanic_class_survival` data.frame counts the number of passengers who survived and those who did not survive within each `class` in the `titanic` data.frame.

Complete the code by filling in the blanks.

```{r}
__BLANK 1__ <- titanic |> 
  count(__BLANK 2__)

```

::: {.callout-tip collapse="true"}
#### Show answer

```{r}
titanic_class_survival <- titanic |> 
  count(class, survived)
```

:::



<br>

### Question 12
How would you describe the variation in the distribution of `age` across `classes` and `genders`?

Complete the code by filling in the blanks.


```{r}
#| echo: false
#| eval: true
#| fig-align: center
#| fig-width: 9
titanic |> 
  ggplot(aes(y = age,
             x = gender,
             fill = gender)) +
  geom_boxplot(show.legend = F) +
  facet_wrap(~class)
```


```{r}
ggplot(data = __BLANK 1__,
       mapping = aes(x = gender,
                     __BLANK 2__ = age,
                     __BLANK 3__ = gender)) +
  __BLANK 4__(show.legend = F) +
  __BLANK 5__(~class) +
  scale_fill_tableau()
```

::: {.callout-tip collapse="true"}
#### Show answer

```{r}
ggplot(data = titanic,
       mapping = aes(x = gender,
                     y = age,
                     fill = gender)) +
  geom_boxplot(show.legend = F) +
  facet_wrap(~class) +
  scale_fill_tableau()
```

:::

<br>

### Question 13
Provide a comment on the variation in the distribution of `age` across `classes` and `genders`.

::: {.callout-tip collapse="true"}
#### Show answer

- For both **female** and **male** groups, the `ages` of the **first class** passengers in the Titanic ranges wider than the **second class** and the **third class**. 
- For both **female** and **male** groups,the median of the **first class** passengers's `ages` is higher than that of the **second class** and the **third class**.
- The first quartile of **female**'s `age` is always lower than that of **male**'s across **all classes**. Particularly, the such gap is wider for the **first class**.

:::

<br>

### Question 14
How would you describe the variation in the distribution of `survived` across `classes` and `genders`?

Complete the code by filling in the blanks.

```{r}
#| echo: false
#| eval: true
#| fig-align: center
#| fig-width: 9

ggplot(titanic,
       aes(y = class,
           fill = survived)) +
  geom_bar() +
  facet_wrap(~gender)
```

```{r}
ggplot(data = __BLANK 1__,
       mapping = aes(__BLANK 2__ = class,
                     __BLANK 3__ = survived)) +
  __BLANK 4__() +
  __BLANK 5__(~gender) +
  labs(x = "Proportion") +
  scale_fill_tableau()
```

::: {.callout-tip collapse="true"}
#### Show answer

```{r}
ggplot(data = titanic,
       mapping = aes(y = class,
                     fill = survived)) +
  geom_bar() +
  facet_wrap(~gender) +
  labs(x = "Proportion") +
  scale_fill_tableau()
```

:::

<br>


### Question 15
How would you describe the variation in the distribution of `survived` across `classes` and `genders`?

Complete the code by filling in the blanks.


```{r}
#| echo: false
#| eval: true
#| fig-align: center
#| fig-width: 9

ggplot(titanic,
       aes(y = class,
           fill = survived)) +
  geom_bar(position = "fill") +
  facet_wrap(~gender) +
  labs(x = "Proportion")
```


```{r}
ggplot(data = __BLANK 1__,
       mapping = aes(__BLANK 2__ = class,
                     __BLANK 3__ = survived)) +
  __BLANK 4__(position = __BLANK 5__) +
  __BLANK 6__(~gender) +
  labs(x = "Proportion") +
  scale_fill_tableau()
```

::: {.callout-tip collapse="true"}
#### Show answer

```{r}
ggplot(data = titanic,
       mapping = aes(y = class,
                     fill = survived)) +
  geom_bar(position = "fill") +
  facet_wrap(~gender) +
  labs(x = "Proportion") +
  scale_fill_tableau()
```


:::

<br>


### Question 16
How would you describe the variation in the distribution of `survived` across `classes` and `genders`?

Complete the code by filling in the blanks.


```{r}
#| echo: false
#| eval: true
#| fig-align: center
#| fig-width: 9

ggplot(titanic,
       aes(y = class,
           fill = survived)) +
  geom_bar(position = "dodge") +
  facet_wrap(~gender) 
```


```{r}
ggplot(data = __BLANK 1__,
       mapping = aes(__BLANK 2__ = class,
                     __BLANK 3__ = survived)) +
  __BLANK 4__(position = __BLANK 5__) +
  __BLANK 6__(~gender) +
  scale_fill_tableau()
```
::: {.callout-tip collapse="true"}
#### Show answer

```{r}
ggplot(data = titanic,
       mapping = aes(y = class,
                     fill = survived)) +
  geom_bar(position = "dodge") +
  facet_wrap(~gender) +
  scale_fill_tableau()
```

:::


<br>


### Question 17
Provide a comment on the variation in the distribution of `survived` across `classes` and `genders`.

::: {.callout-tip collapse="true"}
#### Show answer

- First-class passengers had the highest survival rate. 
  - Survival rates decline progressively in second and third classes.

-	Female passengers generally had a much higher survival rate compared to male passengers.
	- Female first-class passengers had the highest survival rates, followed by female second-class and third-class passengers. 
	- Male survival rates were considerably lower in all classes, with third-class males experiencing the lowest survival likelihood.

- These patterns may be attributed to the influence of both socioeconomic status and gender norms prevalent in the early 1900s.

:::

<br>

## Questions 18-20
Consider the following `nyc_dogs` data.frame for Questions 18-20:

```{r}
#| eval: true
nyc_dogs <- read_csv("https://bcdanl.github.io/data/nyc_dogs_cleaned.csv")
```

```{r}
#| echo: false
#| eval: true
rmarkdown::paged_table(nyc_dogs)
```

- The `nyc_dogs` data.frame contains data on licensed dogs in New York city.



<br>

### Question 18

How would you create the following data.frame, `nyc_dogs_breeds`?

```{r}
#| echo: false
#| eval: true
#| results: asis

nyc_dogs_breeds <- nyc_dogs |> 
  count(breed) |> 
  filter(!is.na(breed)) |> 
  filter(n >= 2000) |> 
  arrange(-n)
rmarkdown::paged_table(nyc_dogs_breeds)
```

- The `nyc_dogs_breeds` data.frame counts the number of occurrences for each value in the `breed` variable in the `nyc_dogs` data.frame.
  - The `nyc_dogs_breeds` data.frame keeps observations if 
    1. The number of occurrences (`n`) is greater than or equal to 2000;
    2. The value of `breed` is not missing.
  - The observations in the `nyc_dogs_breeds` data.frame is arranged by `n` in descending order.


Complete the code by filling in the blanks.

```{r}

__BLANK 1__ <- nyc_dogs |> 
  __BLANK 2__ |> 
  filter(__BLANK 3__(breed)) |> 
  filter(__BLANK 4__) |> 
  arrange(__BLANK 5__)

```



::: {.callout-tip collapse="true"}
#### Show answer

```{r}
nyc_dogs_breeds <- nyc_dogs |> 
  count(breed) |> 
  filter(!is.na(breed)) |> 
  filter(n >= 2000) |> 
  arrange(-n)  # or arrange(desc(n))
```

:::

<br>

### Question 19
How would you describe the distribution of `breed` using the `nyc_dogs_breeds` data.frame?

Complete the code by filling in the blanks.

```{r}
#| echo: false
#| eval: true
#| fig-align: center
#| fig-width: 9
#| fig-height: 15
ggplot(data = nyc_dogs_breeds,
       mapping = aes(x = n,
                     y = breed)) +
  geom_col()
```


```{r}
ggplot(data = __BLANK 1__,
       mapping = aes(x = __BLANK 1__,
                     __BLANK 3__)) +
  __BLANK 4__()
```

::: {.callout-tip collapse="true"}
#### Show answer

```{r}
ggplot(data = nyc_dogs_breeds,
       mapping = aes(x = n,
                     y = breed)) +
  geom_col()
```

:::



<br>

### Question 20
How would you describe the distribution of `breed` using the `nyc_dogs_breeds` data.frame?

Complete the code by filling in the blanks.
```{r}
#| echo: false
#| eval: true
#| fig-align: center
#| fig-width: 9
#| fig-height: 15
ggplot(data = nyc_dogs_breeds,
       mapping = aes(x = n,
                     y = fct_reorder(breed, n))) +
  geom_col() +
  labs(y = "Breed")
```


```{r}
ggplot(data = __BLANK 1__,
       mapping = aes(x = __BLANK 1__,
                     __BLANK 3__)) +
  __BLANK 4__() +
  labs(y = "Breed")
```


::: {.callout-tip collapse="true"}
#### Show answer

```{r}
ggplot(data = nyc_dogs_breeds,
       mapping = aes(x = n,
                     y = fct_reorder(breed, n))) +
  geom_col() +
  labs(y = "Breed")
```

:::

<br>

