---
title: Time Trend Plots
subtitle: Classwork 11
date: last-modified
from: markdown+emoji

execute: 
  eval: true
  echo: true
---

```{r}
#| include: false

library(gapminder)
library(tidyverse)
library(lubridate)
library(skimr)
library(ggthemes)
library(hrbrthemes)
library(viridis)
library(rmarkdown)

theme_set(theme_fivethirtyeight() +
            theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.75),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 0,
                               size = rel(1.75),
                               margin = margin(0,10,0,0)),
                axis.text.x = element_text(size = rel(1.75)),
                axis.text.y = element_text(size = rel(1.75)),
                strip.text = element_text(size = rel(1.5)),
                legend.position = "top",
                legend.text = element_text(size = rel(1.5)),
                legend.title = element_text(size = rel(1.5))
                )
          )


# Set global options for color-blind-friendly scales
scale_colour_discrete <- function(...) scale_colour_viridis_d(...)
scale_fill_discrete <- function(...) scale_fill_viridis_d(...)

```

# Trend in GDP per capita

The `gapminder` package provides the data.frame object, `gapminder`. Let's assign this to `df_gapminder`:

```{r}
library(gapminder)
df_gapminder <- gapminder::gapminder
```


```{r}
#| echo: false

DT::datatable(df_gapminder, options = list(
  pageLength = 12
))
```


## Q1a
- Provide `ggplot()` code to describe the time trend of GDP per capita (`gdpPercap`).
  - How would you take into account country-level data structure?

<br>


## Q1b
- Provide `ggplot()` code to describe how the *overall* time trend of GDP per capita (`gdpPercap`) varies by `continent`.


<br>
