---
title: Color vs. Facet
subtitle: Classwork 10
date: 2024-11-05
from: markdown+emoji

execute: 
  eval: true
  echo: true
---

```{r}
#| include: false

library(gapminder)
library(tidyverse)
library(lubridate)
library(skimr)
library(ggthemes)
library(hrbrthemes)
library(viridis)
library(rmarkdown)

theme_set(theme_fivethirtyeight() +
            theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.75),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 0,
                               size = rel(1.75),
                               margin = margin(0,10,0,0)),
                axis.text.x = element_text(size = rel(1.75)),
                axis.text.y = element_text(size = rel(1.75)),
                strip.text = element_text(size = rel(1.5)),
                legend.position = "top",
                legend.text = element_text(size = rel(1.5)),
                legend.title = element_text(size = rel(1.5))
                )
          )


# Set global options for color-blind-friendly scales
scale_colour_discrete <- function(...) scale_colour_viridis_d(...)
scale_fill_discrete <- function(...) scale_fill_viridis_d(...)

```

# NBC Show Data

The `nbc_show` data is from NBC’s TV pilots:

```{r}
library(tidyverse)
nbc_show <- read_csv("https://bcdanl.github.io/data/nbc_show.csv")
```

- Gross Ratings Points (`GRP`): estimated total viewership of the show, which measures broadcast marketability.
  - A higher GRP suggests broader exposure and a potentially more marketable show.

- Projected Engagement (`PE`): how engaged viewers were after watching a show, a more suitable measure of audience.
  - After watching a show, a viewer is quizzed on order and detail.
  - This measures their engagement with the show (and ads!).
  - High PE values suggest strong engagement.

```{r}
#| echo: false

DT::datatable(nbc_show, options = list(
  pageLength = 40
))
```


## Q1a
- Provide `ggplot()` code to describe the relationship between `GRP` and `PE`.

<br>


## Q1b
- Provide `ggplot()` code to describe how the relationship between `GRP` and `PE` varies by `Genre`.

<br>

## Q1c
- What are the advantages to using faceting instead of the `color` aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?

<br>
