---
title: Classwork 14
subtitle: "Distribution Plots and Counting"
date: 2025-11-17
from: markdown+emoji

fig-width: 10
execute: 
  eval: true
  echo: true
---

```{r}
#| include: false

library(gapminder)
library(tidyverse)
library(lubridate)
library(skimr)
library(ggthemes)
library(hrbrthemes)
library(viridis)
library(rmarkdown)

theme_set(theme_fivethirtyeight() +
            theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.75),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 0,
                               size = rel(1.75),
                               margin = margin(0,10,0,0)),
                axis.text.x = element_text(size = rel(1.75)),
                axis.text.y = element_text(size = rel(1.75)),
                strip.text = element_text(size = rel(1.5)),
                legend.position = "top",
                legend.text = element_text(size = rel(1.5)),
                legend.title = element_text(size = rel(1.5))
                )
          )


# Set global options for color-blind-friendly scales
scale_colour_discrete <- function(...) scale_colour_viridis_d(...)
scale_fill_discrete <- function(...) scale_fill_viridis_d(...)

```


# R Packages

For Classwork 14, please load the following R packages and create the data.frame `nycflights13::flights`:
```{r}
# install.packages("ggthemes")
library(ggthemes)
library(tidyverse)

flights <- nycflights13::flights
```


# Question 1
🤖 **Task:** Fill in the blanks in the provided `ggplot()` code chunk to visualize the distribution of `air_time` (minutes spent in the air).


```{r}
#| echo: false
ggplot(data = flights,
       mapping = aes(x = air_time)) +
  geom_histogram(binwidth = 1)
```

```{r}
#| eval: false

ggplot(data = flights,
       mapping = aes(__BLANK_1__)) +
  geom___BLANK_2__(__BLANK_3__ = 1)
```


<br>


# Question 2
🤖 **Task:** Fill in the blanks in the provided `ggplot()` code chunk to visualize how the distribution of `air_time` (minutes spent in the air) varies by `origin`.

## Part A. Histograms

```{r}
#| echo: false
ggplot(data = flights,
       mapping = aes(x = air_time)) +
  geom_histogram(bins = 50,
                 fill = "lightblue") +
  facet_wrap(~origin)
```

```{r}
#| eval: false
ggplot(data = flights,
       mapping = aes(__BLANK_1__)) +
  geom___BLANK_2__(__BLANK_3__ = 50,
                   __BLANK_4__ = "lightblue") +
  facet_wrap(__BLANK_5__)
```

<br>

## Part B. Boxplots

```{r}
#| echo: false
ggplot(data = flights,
       mapping = aes(x = air_time,
                     y = origin,
                     fill = origin)) +
  geom_boxplot(show.legend = FALSE)
```

```{r}
#| eval: false
ggplot(data = flights,
       mapping = aes(__BLANK_1__,
                     y = __BLANK_2__,
                     __BLANK_3__,
                     )) +
  geom___BLANK_4__(show.legend = FALSE)
```


<br>

# Question 3

🤖 **Task:** Create the data frame `top3_n`, containing **two variables** and **five observations**:

- `carrier`: the **top 3 carriers** ranked by number of flights  
- `n`: the **number of flights** operated by each of those carriers


```{r}
#| echo: false
top3_n <- flights |> 
  count(carrier) |> 
  arrange(-n) |> 
  head(3)

rmarkdown::paged_table(top3_n)
```

```{r}
#| eval: false
__BLANK_1__ <- flights |> 
  __BLANK_2__ |> 
  __BLANK_3__( desc(n) ) |> 
  head(3)  # returns the first 3 observations in the given data.frame
```




<br>


# Question 4
🤖 **Task:** Create the data.frame `top3_carriers`, containing **six variables** (`month`, `day`, `dep_time`, `carrier`, `origin`, and `dest`) and **all observations** for **flights** operated by **only the top 3 carriers** identified in *Question 3*.


```{r}
#| echo: false
top3_carriers <- flights |> 
  filter(carrier == "UA" |
         carrier == "B6" |
         carrier == "EV" ) |> 
  select(month, day, dep_time, carrier, origin, dest)

rmarkdown::paged_table(top3_carriers)
```


```{r}
#| eval: false
__BLANK_1__ <- flights |> 
  filter(__BLANK_2__) |> 
  __BLANK_3__(month, day, dep_time, carrier, origin, dest)
```


<br>


# Question 5
🤖 **Task:** Fill in the blanks in the provided `ggplot()` code chunks to visualize the distribution of `carrier` using the `top3_carriers` data.frame.


## Part A. Bar Charts
```{r}
#| echo: false
ggplot(data = top3_carriers,
       mapping = aes(y = carrier,
                     fill = carrier)) +
  geom_bar(show.legend = FALSE)
```


```{r}
#| eval: false
ggplot(data = top3_carriers,
       mapping = aes(__BLANK_1__,
                     fill = __BLANK_2__)) +
  geom___BLANK_3__(show.legend = FALSE)
```



## Part B. Proportion Bar Charts
```{r}
#| echo: false
ggplot(data = top3_carriers,
       mapping = aes(y = carrier,
                     x = after_stat(prop),
                     group = 1)) +
  geom_bar(show.legend = FALSE)
```


```{r}
#| eval: false
ggplot(data = top3_carriers,
       mapping = aes(__BLANK_1__,
                     x = __BLANK_2__,
                     __BLANK_3__ = 1)) +
  geom___BLANK_4__(show.legend = FALSE)
```



<br>

# Question 6
🤖 **Task:** Fill in the blanks in the provided `ggplot()` code chunks to visualize how the distribution of `carrier` varies by `origin` using the `top3_carriers` data.frame.


## Part A. Stacked Bar Charts

```{r}
#| echo: false
ggplot(data = top3_carriers,
       mapping = aes(y = origin,
                     fill = carrier)) +
  geom_bar()
```


```{r}
#| eval: false
ggplot(data = __BLANK_1__,
       mapping = aes(y = __BLANK_2__,
                     __BLANK_3__)) +
  geom_bar()
```

<br>

## Part B. 100% Stacked Bar Charts
```{r}
#| echo: false

ggplot(data = top3_carriers,
       mapping = aes(y = origin,
                     fill = carrier)) +
  geom_bar(position = "fill") +
  labs(x = "Proportion") # label x-axis title
```

```{r}
#| eval: false
ggplot(data = __BLANK_1__,
       mapping = aes(y = __BLANK_2__,
                     __BLANK_3__)) +
  geom_bar(position = __BLANK_4__) +
  labs(x = "Proportion") # label x-axis title
```


<br>


## Part C. Clustered Bar Charts
```{r}
#| echo: false
ggplot(data = top3_carriers,
       mapping = aes(y = origin,
                     fill = carrier)) +
  geom_bar(position = "dodge")
```

```{r}
#| eval: false
ggplot(data = __BLANK_1__,
       mapping = aes(y = __BLANK_2__,
                     __BLANK_3__)) +
  geom_bar(position = __BLANK_4__)
```


## Part D. Facetted Bar Charts

```{r}
#| echo: false
ggplot(data = top3_carriers,
       mapping = aes(y = carrier,
                     fill = carrier)) +
  geom_bar(show.legend = F) +
  facet_wrap(~origin) +
  theme(
    axis.text.x = element_text(angle = 45,
                               margin = margin(20,0,0,0))
    )
  
```

```{r}
#| eval: false
ggplot(data = __BLANK_1__,
       mapping = aes(y = __BLANK_2__,
                     __BLANK_3__)) +
  geom_bar(show.legend = FALSE) +
  facet_wrap(__BLANK_4__)
```


<br>

# Question 7
🤖 **Task:** Fill in the blanks in the provided `ggplot()` code chunk to visualize the distribution of `carrier` using the `top3_n` data.frame.

```{r}
#| echo: false
ggplot(data = top3_n,
       mapping = aes(x = n,
                     y = carrier,
                     fill = carrier)) +
  geom_col(show.legend = FALSE)
```

```{r}
#| eval: false
ggplot(data = top3_n,
       mapping = aes(x = __BLANK_1__,
                     y = __BLANK_2__,
                     __BLANK_3__)) +
  geom___BLANK_4__(show.legend = FALSE)
```


<br>


# Question 8
🤖 **Task:** Fill in the blanks in the provided `ggplot()` code chunk to visualize the **sorted** distribution of  `carrier`' using the `top3_n` data.frame.

```{r}
#| echo: false
ggplot(data = top3_n,
       mapping = aes(x = n,
                     y = fct_reorder(carrier, n))) +
  geom_col() +
  labs(y = "Carrier")
```


```{r}
#| eval: false
ggplot(data = top3_n,
       mapping = aes(x = __BLANK_1__,
                     y = __BLANK_2__)) +
  geom___BLANK_3__() +
  labs(y = "Carrier")   # label y-axis title
```



<br>



# Question 9

🤖 **Task:** Create a data.frame named `carrier_per_origin` with the following variables:

- `origin`: the origin airport  
- `carrier`: the airline carrier  
- `n`: the number of flights operated by each **carrier** from each **origin** airport  

The `carrier_per_origin` data.frame should contain the **count of flights** for every carrier–origin combination.



```{r}
#| echo: false
carrier_per_origin <- flights |> 
  count(origin, carrier) |> 
  arrange(origin, -n)

rmarkdown::paged_table(carrier_per_origin, 
                       options = list(rows.print = 35))
```


```{r}
#| eval: false

carrier_per_origin <- flights |> 
  __BLANK__ |> 
  arrange(origin, -n)
```



<br>

