---
title: Classwork 7
subtitle: Taxonomy of Data
date: 2025-10-17
execute: 
  echo: false
  eval: true
  warning: false
  message: false
---
```{r}
#| include: false

library(knitr)
library(rmarkdown)
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)
library(DT)

theme_set(theme_ipsum() +
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = element_text(size = rel(1.5)),
                axis.title.y = element_text(size = rel(1.5)),
                legend.title = element_text(size=rel(1.25))
                ))
```

<br>

Consider the following data.frame, `netflix_data`, displayed below:
```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false

netflix_data <- data.frame(
  UserID = 1:10,
  Age = c(25, 34, 28, 45, 23, 37, 31, 29, 41, 26),                   # Ratio
  Gender = c("Female", "Male", "Female", "Male", "Female", "Male", "Female", "Male", "Male", "Female"),  # Nominal
  SubscriptionPlan = factor(c("Basic", "Premium", "Standard", "Premium", "Basic", "Standard", "Premium", "Basic", "Standard", "Premium"),
                            levels = c("Basic", "Standard", "Premium"), ordered = TRUE),                 # Ordinal
  FavoriteGenre = c("Drama", "Comedy", "Action", "Horror", "Sci-Fi", "Romance", "Documentary", "Thriller", "Animation", "Fantasy"),  # Nominal
  HoursWatched = c(15.5, 20.3, 12.0, 25.7, 8.2, 18.9, 22.5, 16.8, 19.4, 14.1),                   # Ratio
  LastLoginTime = c(23.5, 8.2, 15.3, 21.0, 2.5, 13.7, 9.8, 18.4, 7.1, 12.6),                            # Interval (hours since midnight)
  AccountMonths = c(12, 24, 6, 36, 4, 18, 30, 9, 15, 21),                                            # Ratio
  Satisfaction = factor(c(5, 4, 3, 5, 2, 4, 5, 3, 4, 5), ordered = TRUE),                         # Ordinal
  nDevices = c(2, 3, 1, 4, 1, 2, 3, 1, 2, 3),                                                    # Ratio
  LastMovieRating = c(7.8, 6.5, 8.2, 5.9, 9.0, 7.1, 8.5, 6.8, 7.6, 8.0),                     # Interval
  nProfiles = c(1, 2, 1, 3, 1, 2, 2, 1, 2, 3),                                                   # Ratio
  Language = c("English", "Spanish", "English", "French", "English", "German", "English", "Spanish", "English", "French")
)

# DT::datatable(netflix_data)

rmarkdown::paged_table(netflix_data)


# knitr::kable(netflix_data |> select(AccountMonths:Language))

```

### Description of Variables in `netflix_data`:

1. `UserID`: Identifier for each user
2. `Age`: Age of the user in years
3. `Gender`: Gender of the user
4. `SubscriptionPlan`: Type of Netflix subscription
5. `FavoriteGenre`: User’s favorite genre
6. `HoursWatched`: Average hours watched per week
7. `LastLoginTime`: Time of last login in hours since midnight
8. `AccountMonths`: Age of the account in months
9. `Satisfaction`: User satisfaction rating (1 to 5 stars)
10.	`nDevices`: Number of devices connected
11.	`LastMovieRating`: Rating of the last watched movie (1.0 to 10.0)
12.	`nProfiles`: Number of profiles on the account
13.	`Language`: User’s preferred language


### Question 1
What type of variable is `FavoriteGenre` in the dataset?

a. Nominal
b. Ordinal
c. Interval
d. Ratio


::: {.callout-tip collapse="true"}
### Show answer

✅ **Answer: a. Nominal** — `FavoriteGenre` represents non-ordered categories.

:::



### Question 2
What type of variable is `SubscriptionPlan` in the dataset?

a. Nominal
b. Ordinal
c. Interval
d. Ratio


::: {.callout-tip collapse="true"}
### Show answer

✅ **Answer: b. Ordinal** — Subscription plans follow a natural order: Basic < Standard < Premium.

:::



### Question 3
What type of variable is `LastLoginTime` in the dataset?

a. Nominal
b. Ordinal
c. Interval
d. Ratio


::: {.callout-tip collapse="true"}
### Show answer
✅ **Answer: c. Interval** — Time since midnight has meaningful differences but no absolute zero.


:::



### Question 4
What type of variable is `Satisfaction` in the dataset?

a. Nominal
b. Ordinal
c. Interval
d. Ratio


::: {.callout-tip collapse="true"}
### Show answer

✅ **Answer: b. Ordinal** — Satisfaction rating is ranked, but the distance between levels is not precise.

:::


<br><br>

# Discussion 

<font size = "5">Welcome to our Classwork 7 Discussion Board! 👋 </font>

This space is designed for you to engage with your classmates about the material covered in Classwork 7.

Whether you are looking to delve deeper into the content, share insights, or have questions about the content, this is the perfect place for you.

If you have any specific questions for Byeong-Hak (**@bcdanl**) or peer classmate (@GitHub-Username) regarding the Classwork 7 materials or need clarification on any points, don't hesitate to ask here. 

Let's collaborate and learn from each other!
