---
title: Classwork 5
subtitle: R Basics II
date: 2025-09-26
# code-fold: true
execute: 
  echo: true
  warning: false
  message: false
---
```{r}
#| include: false

library(knitr)
library(rmarkdown)
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)

theme_set(theme_ipsum() +
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = element_text(size = rel(1.5)),
                axis.title.y = element_text(size = rel(1.5)),
                legend.title = element_text(size=rel(1.25))
                ))
```



## Question 1.
- Write an R code to calculate the standard deviation (SD) of the integer vector `x` below manually. That is to calculate the SD without using the `sd()` or the `var()` functions.

```{r}
x <- 1:25
```

- Also, write an R code to test whether the standard deviation you calculate manually above is equal to `sd(x)`.


**Answer**:
```{r}
```


<br><br>



## Question 2.
- Write an R code to calculate the range (difference between the maximum and minimum values) of the vector `temp <- c(22, 28, 31, 25, 29)`.

```{r}
temp <- c(22, 28, 31, 25, 29)
```

**Answer**:
```{r}
```

<br><br>


## Question 3.
- Consider the vectors:

```{r}
my_vec <- c(-10, -20, 30, 10, 50, 40, -100)
beers <- c("BUD LIGHT", "BUSCH LIGHT", "COORS LIGHT", 
           "GENESEE LIGHT", "MILLER LITE", "NATURAL LIGHT")
```


- Write an R code to filter only the positive values in `my_vec`.
- Write an R code to access the beers that are in positions 2, 4, and 6 using indexing.

**Answer**:
```{r}

```


<br><br>


## Question 4.
- Create a logical vector `logical_vec` that checks whether the elements of the vector `ages <- c(15, 22, 18, 24, 30)` are greater than or equal to 20.

```{r}
ages <- c(15, 22, 18, 24, 30)
```

**Answer**:
```{r}
```


<br><br>
