---
title: Classwork 12
subtitle: Color vs. Facet
date: 2025-11-10
from: markdown+emoji

fig-width: 10
execute: 
  eval: true
  echo: true
---

```{r}
#| include: false

library(gapminder)
library(tidyverse)
library(lubridate)
library(skimr)
library(ggthemes)
library(hrbrthemes)
library(viridis)
library(rmarkdown)

theme_set(theme_fivethirtyeight() +
            theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.75),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 0,
                               size = rel(1.75),
                               margin = margin(0,10,0,0)),
                axis.text.x = element_text(size = rel(1.75)),
                axis.text.y = element_text(size = rel(1.75)),
                strip.text = element_text(size = rel(1.5)),
                legend.position = "top",
                legend.text = element_text(size = rel(1.5)),
                legend.title = element_text(size = rel(1.5))
                )
          )


# Set global options for color-blind-friendly scales
scale_colour_discrete <- function(...) scale_color_tableau(...)
scale_fill_discrete <- function(...) scale_fill_tableau(...)

```


# R Packages

For Classwork 12, please load the `tidyverse` package:
```{r}
library(tidyverse)
```


# Question 1. NBC Show Data

The `nbc_show` dataset comes from **NBC’s TV pilots**, containing information about television shows, their **viewership metrics**, and **audience engagement**.

```{r}
nbc_show <- read_csv("https://bcdanl.github.io/data/nbc_show.csv")
```

- **Gross Ratings Points** (`GRP`):  
  Measures the **estimated total viewership** of a show — an indicator of its **broadcast marketability**.  
  - 📺 A higher `GRP` suggests **broader exposure** and a more **marketable** program.

- **Projected Engagement** (`PE`):  
  Captures how **attentive and engaged** viewers were after watching a show — a more suitable measure of **audience engagement**.  
  - 🧠 After viewing, audiences take a short quiz testing **order and detail recall**.  
  - This reflects their **level of attention and retention** (for both the show *and* its ads).  
  - High `PE` values indicate **strong viewer engagement**.

```{r}
#| echo: false

DT::datatable(nbc_show, options = list(
  pageLength = 40
))
```



## Tasks
- 🤖 **Task 1:** Fill in the blanks in the provided `ggplot()` code chunk.  
- 💬 **Task 2:** Add a brief comment describing **the relationship between gross ratings points (`GRP`) and projected engagement (`PE`) varies by genre (`Genre`)**.




### (1) Color
```{r}
#| echo: false
ggplot(data = nbc_show,
       mapping = aes(x = GRP,
                     y = PE,
                     color = Genre)) +
  geom_point() +
  geom_smooth(method = "lm",
              se = FALSE)   # se = FALSE turns off the ribbon
```

```{r}
#| eval: false
ggplot(__BLANK_1__ = nbc_show,
       mapping = aes(x = GRP,
                     y = PE,
                     __BLANK_2__ = Genre)) +
  geom_point() +
  geom_smooth(__BLANK_3__,
              se = FALSE)   # se = FALSE turns off the ribbon
```

<br>

### (2) Facet
```{r}
#| echo: false
ggplot(data = nbc_show,
       mapping = aes(x = GRP,
                     y = PE)) +
  geom_point() + 
  geom_smooth(method = "lm",
              se = FALSE) +  # se = FALSE turns off the ribbon
  facet_wrap(~Genre)
```

```{r}
#| eval: false
ggplot(data = nbc_show,
       mapping = aes(x = GRP,
                     y = PE)) +
  geom_point() +  
  geom_smooth(method = __BLANK_1__,
              se = FALSE) +  # se = FALSE turns off the ribbon
  __BLANK_2___wrap(__BLANK_3__)
```

<br>

### (3) Facet with Color
```{r}
#| echo: false
ggplot(data = nbc_show,
       mapping = aes(x = GRP,
                     y = PE,
                     color = Genre)) +
  geom_point(show.legend = FALSE) + 
  geom_smooth(method = "lm",
              show.legend = FALSE, # show.legend = FALSE turns of legend
              se = FALSE) +  # se = FALSE turns off the ribbon
  facet_wrap(~Genre)
```

```{r}
#| eval: false
ggplot(data = nbc_show,
       mapping = aes(x = GRP,
                     y = PE,
                     color = __BLANK_1__)) +
  geom_point(show.legend = FALSE) +  # show.legend = FALSE turns of legend
  geom_smooth(method = __BLANK_2__,
              show.legend = FALSE,   # show.legend = FALSE turns of legend
              se = FALSE) +  # se = FALSE turns off the ribbon
  __BLANK_3___wrap(__BLANK_4__)
```


<br>

# Question 2. GDP per capita and Life Expectancy 

For **Question 2**, please **load the R package `gapminder`** before starting:
```{r}
#| eval: false

# install.packages("gapminder")
library(gapminder)
??gapminder
```

The **`gapminder`** package provides a built-in dataset named **`gapminder`**, which contains country-level data on life expectancy, GDP per capita, and population across time.  

Let's assign it to a new object called **`df_gapminder`**:

```{r}
df_gapminder <- gapminder::gapminder
```


```{r}
#| echo: false

DT::datatable(df_gapminder, options = list(
  pageLength = 12
))
```


## Tasks
- 🤖 **Task 1:** Fill in the blanks in the provided `ggplot()` code chunk.  
- 💬 **Task 2:** Add a brief comment describing **the relationship between GDP per capita (`gdpPercap`) and life expectancy (`lifeExp`) varies by continents (`continent`)**.



### (1) Color: Only Scatterplot
```{r}
#| echo: false
ggplot(data = df_gapminder,
       mapping = aes(x = log(gdpPercap),
                     y = lifeExp,
                     color = continent)) +  # different colors are used to distinguish continents
  geom_point(alpha = .5)  # Add 50% transparency to reduce overplotting
```

```{r}
#| eval: false
ggplot(__BLANK_1__ = df_gapminder,
       mapping = aes(__BLANK_2__ = log(gdpPercap),
                     __BLANK_3__ = lifeExp,
                     __BLANK_4__ = continent)) +  # different colors are used to distinguish continents
  geom_point(__BLANK_5__)  # Add 50% transparency to reduce overplotting
```

<!-- - While transparency (`alpha`) in the scatterplot partially reduces overplotting, it does not fully address the issue, especially in dense regions.  -->
<!-- - This is because, in general, the mixing of overlapping transparent colors may be no longer represent the colors of the categories.  -->
<!-- - Adding fitted lines clarifies the differences in relationships across continents. -->

<br>

### (2) Color: Scatterplot with Fitted Line
```{r}
#| echo: false
ggplot(data = df_gapminder,
       mapping = aes(x = log(gdpPercap),
                     y = lifeExp,
                     color = continent)) +  # different colors are used to distinguish continents
  geom_point(alpha = .5) + # Add transparency to reduce overplotting
  geom_smooth(method = "lm")
```

```{r}
#| eval: false
ggplot(__BLANK_1__ = df_gapminder,
       mapping = aes(__BLANK_2__ = log(gdpPercap),
                     __BLANK_3__ = lifeExp,
                     __BLANK_4__ = continent)) +  # different colors are used to distinguish continents
  geom_point(__BLANK_5__) +  # Add 50% transparency to reduce overplotting
  geom___BLANK_6__(method = "lm")
```

<!-- - The different slopes of the fitted lines across continents imply that the relationship between GDP per capita and life expectancy differs by continent.  -->
<!--   - Continents like the Americas and Oceania display steeper slopes, indicating a stronger positive association between GDP per capita and life expectancy.  -->
<!--   - This suggests that for the same percentage increase in GDP per capita, the improvement in life expectancy is greater in these regions compared to others. -->

<br>

### (3) Facet: Scatterplot with Fitted Line
```{r}
#| echo: false
ggplot(data = df_gapminder,
       mapping = aes(x = log(gdpPercap),
                     y = lifeExp,
                     color = continent)) +
  geom_point(alpha = .3) +
  geom_smooth(method = "lm") +
  facet_wrap(~continent)
```

```{r}
#| eval: false
ggplot(__BLANK_1__ = df_gapminder,
       mapping = aes(__BLANK_2__ = log(gdpPercap),
                     __BLANK_3__ = lifeExp,
                     __BLANK_4__ = continent)) +
  geom_point(__BLANK_5__ = 0.3) +  # Add 70% transparency to reduce overplotting
  geom___BLANK_6__(method = "lm") +
  facet___BLANK_7__(~continent)
```

<!-- - The faceted view significantly reduces overplotting and provides a more detailed look at regional differences. -->
<!--   - However, using `color` only can make it easier to compare the slope of the fitted lines across continents. -->


<br>


# Question 3. Color vs. Facet
- What are the **advantages** of using **faceting** instead of the `color` aesthetic?  
- What are the **disadvantages**?  
- How might this **trade-off** change if you were working with a **larger dataset**?



<!-- - Advantages of Faceting -->
<!-- 	1. Clarity for Multiple Categories: Faceting avoids the visual clutter of overlapping points and lines, especially when the number of categories (e.g., `Genre`) is large. -->
<!-- 	2. Highlights Individual Patterns: By splitting the data into separate plots, it’s easier to observe specific trends or outliers within each genre. -->
<!-- 	3. Improved Readability: Each genre gets its own visual space, avoiding the need to distinguish between multiple colors. -->

<!-- - Disadvantages of Faceting -->
<!-- 	- Difficult Cross-Category Comparison: Observations in separate facets cannot be directly compared. Audiences need to read through all facets. -->

<!-- - Impact with Larger Dataset -->
<!-- 	-	More Data Points: Overlapping increases, making faceting a more practical option to reduce clutter. Color aesthetics may struggle to show patterns with dense data points. -->
<!-- 	-	More Categories: Differentiating between colors becomes harder as the number of categories increases. In such cases, faceting is clearer. -->
<!-- 	-	Transparency Issues: Using transparency (`alpha`) with many overlapping points and colors can result in a loss of clear category identification. -->

<br>
