---
title: Distribution Plots
subtitle: Classwork 12
date: 2024-11-11
from: markdown+emoji

fig-width: 10
execute: 
  eval: true
  echo: true
---

```{r}
#| include: false

library(gapminder)
library(tidyverse)
library(lubridate)
library(skimr)
library(ggthemes)
library(hrbrthemes)
library(viridis)
library(rmarkdown)
library(nycflights13)

theme_set(theme_fivethirtyeight() +
            theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.75),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 0,
                               size = rel(1.75),
                               margin = margin(0,10,0,0)),
                axis.text.x = element_text(size = rel(1.75)),
                axis.text.y = element_text(size = rel(1.75)),
                strip.text = element_text(size = rel(1.5)),
                legend.position = "top",
                legend.text = element_text(size = rel(1.5)),
                legend.title = element_text(size = rel(1.5))
                )
          )


# Set global options for color-blind-friendly scales
scale_colour_discrete <- function(...) scale_colour_viridis_d(...)
scale_fill_discrete <- function(...) scale_fill_viridis_d(...)

```


For this classwork, consider the `flights` data.frame

```{r}
#| eval: true
library(tidyverse)
library(ggthemes)
library(nycflights13)

flights <- flights
```


```{r}
#| echo: false

rmarkdown::paged_table(flights)
```


## Description of Variables in `flights`

Use `??flights` to see the description of variables in the `flights` data.frame.

<br>

## Q1a
- Provide `ggplot()` code to describe the distribution of `air_time` (amount of time spent in the air, in minutes).

```{r}
#| eval: false

ggplot(data = __BLANK 1__,
       mapping = aes(__BLANK 2__)) +
  __BLANK 3__()
```

**Answer:**

- Since the larger `air_time` means the longer flight, this integer variable could be numeric.
  - `geom_histogram()` would be a safer choice.

```{r}
ggplot(data = flights,
       mapping = aes(x = air_time)) +
  geom_histogram()
```

- In ggplot, the distribution of an integer variable can appear quite similar when using `geom_bar()` and `geom_histogram()`.

```{r}
ggplot(data = flights,
       mapping = aes(x = air_time)) +
  geom_histogram(binwidth = 1)
```

```{r}
ggplot(data = flights,
       mapping = aes(x = air_time)) +
  geom_bar()
```

- In Python or other visualization tools, they can be quite different.

<br>


## Q1b
- Provide `ggplot()` code to describe how the distribution of `air_time` varies by `origin`.

```{r}
#| eval: false
ggplot(data = __BLANK 1__,
       mapping = aes(__BLANK 2__)) +
  __BLANK 3__() +
  facet_wrap(__BLANK 4__)
```

**Answer:**

```{r}
ggplot(data = flights,
       mapping = aes(x = air_time)) +
  geom_histogram(bins = 60) +
  facet_wrap(~origin)
```



<br>



## Q1c
- Create the data.frame, `top5_n`, which include the two variables and 5 observations:
  - `carrier`: the value of the top 5 carriers in terms of the number of flights.
  - `n`: the number of flights operated by each of the top 5 carriers.

```{r}
#| echo: false
#| eval: true
top5_n <- flights |> 
  count(carrier) |> 
  arrange(-n) |> 
  head(5)

rmarkdown::paged_table(top5_n)
```


```{r}
#| eval: false
top5_n <- flights |> 
  __BLANK__ |> 
  arrange(-n) |> 
  head(5)  # returns the first 5 observations of the new data.frame
```

**Answer:**

```{r}
top5_n <- flights |> 
  count(carrier) |> 
  arrange(-n) |> 
  head(5)
```


<br>


## Q1d
- Create the data.frame, `top5_carriers`, which includes all the flights operated by only the top 5 carriers found in Q1c.

```{r}
#| eval: false
top5_carriers <- flights |> 
  filter(carrier == "__BLANK 1__" |
         carrier == "__BLANK 2__" |
         carrier == "__BLANK 3__" |
         carrier == "__BLANK 4__" |
         carrier == "__BLANK 5__" ) 
```

**Answer:**

- We can see which carriers are the top 5 from the `top5_n`:
```{r}
top5_carriers <- flights |> 
  filter(carrier == "UA" |
         carrier == "B6" |
         carrier == "EV" |
         carrier == "DL" |
         carrier == "AA" ) 
```



<br>


## Q1e
- Provide `ggplot()` code to describe the distribution of `carrier` using the `top5_carriers` data.frame.

```{r}
#| eval: false
ggplot(data = __BLANK 1__,
       mapping = aes(__BLANK 2__)) +
  __BLANK 3__()
```

**Answer:**

- Here we are using a horizontal bar chart:
```{r}
ggplot(data = top5_carriers,
       mapping = aes(y = carrier)) +
  geom_bar()
```




<br>

## Q1f
- Provide `ggplot()` code to describe how the distribution of `carrier` varies by `origin` using the `top5_carriers` data.frame.

### Stacked Bar Chart
```{r}
#| eval: false
ggplot(data = __BLANK 1__,
       mapping = aes(y = __BLANK 2__,
                     __BLANK 3__)) +
  geom_bar()
```

**Answer:**

```{r}
ggplot(data = top5_carriers,
       mapping = aes(y = origin,
                     fill = carrier)) +
  geom_bar()
```



### 100% Stacked Bar Chart
```{r}
#| eval: false
ggplot(data = __BLANK 1__,
       mapping = aes(y = __BLANK 2__,
                     __BLANK 3__)) +
  geom_bar(position = __BLANK 4__) +
  labs(x = "Proportion")
```

**Answer:**

```{r}
ggplot(data = top5_carriers,
       mapping = aes(y = origin,
                     fill = carrier)) +
  geom_bar(position = "fill")
```



### Clustered Bar Chart
```{r}
#| eval: false
ggplot(data = __BLANK 1__,
       mapping = aes(y = __BLANK 2__,
                     __BLANK 3__)) +
  geom_bar(position = __BLANK 4__)
```


```{r}
ggplot(data = top5_carriers,
       mapping = aes(y = origin,
                     fill = carrier)) +
  geom_bar(position = "dodge")
```

- c.f., `dodge2` provides a gap between bars:
```{r}
ggplot(data = top5_carriers,
       mapping = aes(y = origin,
                     fill = carrier)) +
  geom_bar(position = "dodge2")
```


### Facetted Bar Chart
```{r}
#| eval: false
ggplot(data = __BLANK 1__,
       mapping = aes(y = __BLANK 2__,
                     __BLANK 3__)) +
  geom_bar(show.legend = F) +
  facet_wrap(__BLANK 4__)
```

**Answer:**
```{r}
#| eval: false
ggplot(data = top5_carriers,
       mapping = aes(y = carrier,
                     fill = carrier)) +
  geom_bar(show.legend = F) +
  facet_wrap(~origin)
```

```{r}
#| echo: false
ggplot(data = top5_carriers,
       mapping = aes(y = carrier,
                     fill = carrier)) +
  geom_bar(show.legend = F) +
  facet_wrap(~origin) +
  theme(
    axis.text.x = element_text(angle = 45,
                               margin = margin(20,0,0,0))
    )
  
```

<br>

## Q1g
- Provide `ggplot()` code to describe the distribution of `carrier` using the `top5_n` data.frame.

```{r}
#| eval: false
ggplot(data = top5_n,
       mapping = aes(x = __BLANK 1__,
                     y = __BLANK 2__)) +
  __BLANK 3__()
```

**Answer:**

```{r}
ggplot(data = top5_n,
       mapping = aes(x = n,
                     y = carrier)) +
  geom_col()
```

<br>


## Q1h
- Provide `ggplot()` code to describe the **sorted** distribution of  `carrier`' using the `top5_n` data.frame.

```{r}
#| eval: false
ggplot(data = top5_n,
       mapping = aes(x = __BLANK 1__,
                     y = __BLANK 2__)) +
  __BLANK 2__()
```

**Answer:**


```{r}
ggplot(data = top5_n,
       mapping = aes(x = n,
                     y = fct_reorder(carrier, n))) +
  geom_col()
```


- `labs()` can label y-axis title.

```{r}
ggplot(data = top5_n,
       mapping = aes(x = n,
                     y = fct_reorder(carrier, n))) +
  geom_col() +
  labs(y = "Carrier")
```


<br>



## Q1i
-	Create the following data frame named `carrier_per_origin` with the following three variables:
	-	`origin`: the origin airport
	-	`carrier`: the airline carrier
	-	`n`: the number of flights operated by each **carrier** from each **origin** airport
-	The `carrier_per_origin` data frame should contain the count of flights operated by each **carrier** departing from each **origin** airport.


```{r}
#| echo: false
#| eval: true
carrier_per_origin <- flights |> 
  count(origin, carrier) |> 
  arrange(origin, -n)

rmarkdown::paged_table(carrier_per_origin)
```


```{r}
#| eval: false

carrier_per_origin <- flights |> 
  __BLANK__ |> 
  arrange(origin, -n)
```

**Answer:**

```{r}
carrier_per_origin <- flights |> 
  count(origin, carrier) |> 
  arrange(origin, -n)
```

<br>

