---
title: Relationship Plots
subtitle: Classwork 9
date: 2024-10-29
from: markdown+emoji

execute: 
  eval: true
  echo: true
---

```{r}
#| include: false

library(gapminder)
library(tidyverse)
library(lubridate)
library(skimr)
library(ggthemes)
library(hrbrthemes)
library(viridis)
library(rmarkdown)

theme_set(theme_fivethirtyeight() +
            theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.75),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 0,
                               size = rel(1.75),
                               margin = margin(0,10,0,0)),
                axis.text.x = element_text(size = rel(1.75)),
                axis.text.y = element_text(size = rel(1.75)),
                strip.text = element_text(size = rel(1.5)),
                legend.position = "top",
                legend.text = element_text(size = rel(1.5)),
                legend.title = element_text(size = rel(1.5))
                )
          )


# Set global options for color-blind-friendly scales
scale_colour_discrete <- function(...) scale_colour_viridis_d(...)
scale_fill_discrete <- function(...) scale_fill_viridis_d(...)

```

# Ice Cream Sales and Drowning Incidents
Consider the data.frame, `df`, the dataset recording monthly ice cream sales and drowning incidents. 

```{r}
df <- read_csv("http://bcdanl.github.io/data/icecream-drowning.csv")
```


```{r}
#| echo: false

DT::datatable(df, options = list(
  pageLength = 12
))
```


## Q1a
- Provide both `ggplot()` and comment to describe the relationship between ice cream sales (`IceCreamSales`) and drowning incidents (`DrowningIncidents`).

<br>

## Q1b
- Is the relationship correlation or causation? Why?

<br><br>



# GDP per capita and Life Expectancy 

For this Section, please install the R package, `gapminder`:
```{r}
#| eval: false

install.packages("gapminder")
??gapminder
```

The `gapminder` package provides the data.frame object, `gapminder`. Let's assign this to `df_gapminder`:

```{r}
df_gapminder <- gapminder::gapminder
```


```{r}
#| echo: false

DT::datatable(df_gapminder, options = list(
  pageLength = 12
))
```


## Q2a
- Provide both `ggplot()` and comment to describe the relationship between GDP per capita (`gdpPercap`) and life expectancy (`lifeExp`).

<br>

## Q2b
- Provide both `ggplot()` and comment to describe how the relationship between GDP per capita (`gdpPercap`) and life expectancy (`lifeExp`) varies by `continent`.

<br>
