---
title: Classwork 6
subtitle: Data Transformation with R
date: 2025-10-01
execute: 
  echo: false
  eval: true
  warning: false
  message: false
---
```{r}
#| include: false

library(knitr)
library(rmarkdown)
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)
library(DT)

theme_set(theme_ipsum() +
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = element_text(size = rel(1.5)),
                axis.title.y = element_text(size = rel(1.5)),
                legend.title = element_text(size=rel(1.25))
                ))
```


# Part 1. `filter()`, `arrange()`, and `distinct()`
For Part 1, consider the `nycflights13::flights` data.frame.

```{r}
#| echo: true

library(tidyverse)
library(nycflights13)
flights <- flights
```

```{r}
#| echo: false

rmarkdown::paged_table(nycflights13::flights)

```

For detailed descriptions of the variables in this data frame, load its **help** documentation using the following command:

```{r}
#| echo: true

# Documentation for the `flights` data.frame from the `nycflights13` package
?flights   # displays the description of variables in the `flights` data.frame

```


## Question 1

- Find all flights that had an arrival delay of two or more hours.
- Find all flights that flew to Houston (**IAH** or **HOU**).
- Find all flights that departed in summer (**July**, **August**, and **September**).
- Find all flights that arrived more than two hours late, but did not leave late.

**Answer**:
```{r}

```



<br><br>

## Question 2
- How many flights have a missing `dep_time`?

**Answer**:
```{r}

```


<br><br>

## Question 3
- Sort flights to find the most delayed flights.

**Answer**:
```{r}

```



<br><br>

## Question 4
- Was there a flight on every day of 2013?

**Answer**:
```{r}

```



<br><br>



# Part 2. `select()` and `rename()`

Consider the following data.frame, `nyc_payroll_new`.

```{r}
#| echo: true

library(tidyverse)
nyc_payroll_new <- read_csv("https://bcdanl.github.io/data/nyc_payroll_2024.csv")
```


```{r}
#| echo: false

rmarkdown::paged_table(nyc_payroll_new)
```

<br>

For detailed descriptions of the variables in this data.frame, please refer to the following link:  

- [Citywide Payroll Data (Fiscal Year)](https://data.cityofnewyork.us/City-Government/Citywide-Payroll-Data-Fiscal-Year-/k397-673e/about_data)

<br>


## Question 1
Create a new data.frame, `df` that keeps only the following five variables---`Fiscal_Year`, `Agency_Name`, `First_Name`, `Last_Name`, and `Base_Salary`---from the data.frame `nyc_payroll_new`.

**Answer**:

```{r}
```


<br><br>

## Question 2
Given the data.frame `df` with a variable named `Agency_Name`, how would you rename it to `Agency`?

**Answer**:

```{r}
```


<br><br>

## Question 3
How would you remove the `Fiscal_Year` variable using `select()`?

**Answer**:

```{r}

```

<br><br>

# Discussion 

<font size = "5">Welcome to our Classwork 6 Discussion Board! 👋 </font>

This space is designed for you to engage with your classmates about the material covered in Classwork 6.

Whether you are looking to delve deeper into the content, share insights, or have questions about the content, this is the perfect place for you.

If you have any specific questions for Byeong-Hak (**@bcdanl**) or peer classmate (@GitHub-Username) regarding the Classwork 6 materials or need clarification on any points, don't hesitate to ask here. 

Let's collaborate and learn from each other!
